### 2024-10-24 _tex/manuscript.tex
Built site for gh-pages

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-mathphys-num,
]{sn-jnl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\usepackage{multirow}
\usepackage{centernot}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Probabilistic Model of Bilateral Lymphatic Spread in Head and
Neck Cancer]{Probabilistic Model of Bilateral Lymphatic Spread in Head
and Neck Cancer}

% author setup
\author*[1,2]{\fnm{Roman} \sur{Ludwig}}\email{roman.ludwig@usz.ch}\author[1,2]{\fnm{Yoel Perez} \sur{Haas}}\email{yoel.perezhaas@usz.ch}\author[1,2]{\fnm{Jan} \sur{Unkelbach}}\email{jan.unkelbach@usz.ch}
% affil setup
\affil[1]{\orgdiv{Department of Physics}, \orgname{University of
Zurich}}
\affil[2]{\orgdiv{Radiation Oncology}, \orgname{University Hospital
Zurich}}

% abstract

\abstract{Purpose: According to current guidelines for elective nodal
irradiation of oropharyngeal squamous cell carcinoma (OPSCC) patients,
large parts of the contralateral lymphatic system are included in the
elective clinical target volume (CTV-N), even for lateralized tumors
without clinical lymph node involvement in the contralateral neck. In
this work, we present a probabilistic model for bilateral lymphatic
tumor progression in OPSCC to predict the personal risk for occult
disease in any lymph node level (LNL), given the patient's clinical
lymph node involvement, T-stage, and lateralization of the primary
tumor.

Methods: We extend a hidden markov model for lymphatic tumor
progression, which was previously developed for ipsilateral lymph node
involvement, to the contralateral neck. The model represents each of the
LNLs I, II, III, IV, V, and VII of both sides of the neck as a hidden
binary random variable with states healthy and involved. LNLs are
connected to the tumor and among each other via arcs that correspond to
spread probabilities. These spread probability rates are learned via
Markov chain Monte Carlo (MCMC) sampling from a dataset of 833 OPSCC
patients.

Results: The model is able to describe the data on lymph node
involvement well with a small number of interpretable parameters.
Midline extension of the primary tumor is the main risk factor for
contralateral involvement. In addition, the risk of occult metastases in
contralateral lymph node level increases with more advanced T-stage and
more severe ipsilateral involvement. The probability of involvement in
contralateral level III is very low if the upstream level II is
clinically negative. Similarly, the probability of involvement in
contralateral level IV is very low if the upstream level III is
clinically negative.

Conclusions: The model may guide personalized volume reduction of the
elective CTV-N. It suggests that the contralateral neck may be spared
from elective irradiation for lateralized tumors not crossing the
midline. In patients with primary tumors crossing the midline but
clinically negative contralateral neck, the contralateral elective
irradiation may be limited to level II.}

% keywords

\begin{document}
\maketitle

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\section{Introduction}\label{introduction}

When treating head and neck squamous cell carcinomas (HNSCC) with
radiotherapy or surgery, not only the primary tumor and the clinically
detected lymph node metastases are targeted. Per current guidelines,
large volumes of the neck are part of the elective clinical target
volume (CTV-N)
\citep{gregoire_ct-based_2003, gregoire_delineation_2014, gregoire_delineation_2018, eisbruch_intensity-modulated_2002, biau_selection_2019, chao_determination_2002, vorwerk_guidelines_2011, ferlito_elective_2009}.
This is to minimize the risk of regional recurrences due to untreated
microscopic disease which cannot be detected by current in-vivo imaging
modalities such as computed tomography (CT), magnetic resonance imaging
(MRI), or positron emission tomography (PET). However, minimizing the
risk of missing occult disease in the lymph drainage region must be
balanced against the toxicity related to unnecessary treatment of
healthy tissue.

The aforementioned guidelines for the definition of the CTV-N are based
on anatomically defined lymph node levels (LNL)
\citep{gregoire_delineation_2014} and the overall prevalence of lymph
node metastases in the LNLs. Based on this, they recommend extensive
irradiation of both sides of the neck for a majority of patients.
However, the overall prevalence of involvement in a given LNL does not
equal a patient's personal risk for occult disease in that LNL, which
depends on an individual patient's state of tumor progression.
Currently, a patient presenting with no clinically detectable nodal
disease (cN0) and a small, clearly lateralized T1 tumor, the same
recommended contralateral CTV-N as a patient with substantial nodal
involvement in the ipsilateral side and an advanced primary tumor that
crosses the mid-sagittal plane. Both patients receive extensive elective
irradiation of the contralateral LNLs II, III, and IVa
\citep{biau_selection_2019}.

To better quantify this personalized risk of occult disease we have
previously developed an intuitive, probabilistic hidden Markov model
\citep{ludwig_hidden_2021, ludwig_modelling_2023} which was originally
based on a conceptually similar Bayesian network model
\citep{pouymayou_bayesian_2019}. So far, these models were limited to
the ipsilateral side of the neck. The main contribution of this work is
the extension of the formalism to include risk predictions for the
contralateral side as well. The model may then guide the reduction of
the electively irradiated volumes in the contralateral neck for patients
whose individual risk for occult disease is low. In turn, this may
substantially reduce the toxicity from irradiation and the associated
decrease in quality of life.

The main contributions of this paper are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  In section section~\ref{sec-data} we present a multi-centric dataset
  on lymph node involvement in 833 OPSCC patients. Based on the dataset
  the main risk factors for contralateral lymph node involvement are
  identified and the requirements for the bilateral model extension are
  described (section~\ref{sec-requirements}).
\item
  In section~\ref{sec-ext-to-contra} we present a bilateral HMM of
  lymphatic progression that accounts for the T-stage lateralization of
  the primary tumor as well as clinical involvement as risk factors for
  occult contralateral involvement. In section~\ref{sec-methods} we
  describe how the model training and the computational experiments were
  set up.
\item
  In the results in section~\ref{sec-results}, we demonstrate that the
  model is able to describe the patterns of contralateral lymph node
  involvement observed in the dataset, and we apply the model to
  estimating risk of occult disease for typical patients. In
  section~\ref{sec-discussion}, implications for volume-deescalated
  radiotherapy are discussed.
\end{enumerate}

\section{Data on Lymphatic Progression Patterns}\label{sec-data}

To be able to create models for lymphatic tumor progression that include
all relevant LNLs including the contralateral side, we have collected a
detailed dataset of 833 patients with newly diagnosed oropharyngeal
squamous cell carcinomas
\citep{ludwig_dataset_2022, ludwig_multi-centric_2023}. It reports the
lymph node involvement per LNL of every patient individually in tabular
form, in addition to other primary tumor and patient characteristics
such as T-category, subsite, lateralization of the primary tumor, and
HPV p16 status. Their patient records have been collected at four
different institutions and a brief overview over some of their patients'
characteristics are shown in table~\ref{tbl-data-overview}. The data
from the Inselspital Bern (ISB) and the Centre Léon Bérard (CLB) only
consist of patients who all received a neck dissection. The majority of
patients from the University Hospital Zürich (USZ) and the Hospital Vall
d'Hebron (HVH) was treated with definitive radiotherapy. Since sugical
treatment is more common for early T-category patients, the ISB and CLB
datasets contain a larger portion of early T-category patients compared
to the USZ and HVH dataset. For a subset of 83 patients in the CLB
dataset, the lateralization of the primary tumor was not reported.

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\begin{longtable}[]{@{}lllllll@{}}

\caption{\label{tbl-data-overview}Overview over the five datasets from
four different institutions used to train and evaluate our model. Here,
we briefly characterize the total number of OPSCC patients from the
respective institution, their median age, what proportion received some
form of neck dissection, the N0 portion of patients, what percentage
presented with early T-category, and the prevalence of primary tumor
midline extension. For a much more detailed look at the data, visit
\href{https://lyprox.org}{lyprox.org}.}

\tabularnewline

\caption{}\label{T_d2cbf}\tabularnewline
\toprule\noalign{}
Institution & Total & Age (median) & Neck Dissection & N0 & Early T-Cat.
& Mid. Ext. \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Institution & Total & Age (median) & Neck Dissection & N0 & Early T-Cat.
& Mid. Ext. \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Centre Léon Bérard & 325 & 60 & 100\% & 19\% & 69\% & 18\% \\
Inselspital Bern & 74 & 61 & 100\% & 18\% & 66\% & 14\% \\
University Hospital Zurich & 287 & 66 & 26\% & 18\% & 52\% & 31\% \\
Vall d\textquotesingle Hebron Barcelona Hospital & 147 & 58 & 5\% & 21\%
& 34\% & 34\% \\

\end{longtable}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\subsection{Consensus on Involvement Status}\label{sec-data-consensus}

Pathological involvement is available only for the subset of surgical
patients and only for the dissected levels. For the remaining patients,
only clinical involvement is available. For the analysis performed in
this paper, all available diagnostic information has been combined into
a consensus decision representing the most likely state of involvement
for each patient and LNL, using literature values for sensitivity and
specificity of different modalities
\citep{de_bondt_detection_2007, kyzas_18f-fluorodeoxyglucose_2008}. How
this consensus is formed is described in detail in
section~\ref{sec-consensus}. In brief, pathology after neck dissection
is taken as the ground truth if available, overwriting any contradicting
clinical involvement diagnoses. If a level was not dissected, the most
likely state is determined by PET-CT in most cases.

\subsection{Data Availability}\label{data-availability}

The entire data, including additional patients with tumors in other
primary locations than the oropharynx, is publicly available: It may be
\href{https://lyprox.org/patients/dataset}{downloaded from LyProX} where
it can be interactively explored too,
\href{https://github.com/rmnldwg/lydata}{from GitHub},
\href{https://zenodo.org/search?q=lydata}{from zenodo}, or via the
\emph{Data-in-Brief} publications \citet{ludwig_dataset_2022} and
\citet{ludwig_multi-centric_2023}. The Data-in-Brief publications
contain a detailed description of the datasets and the data format.
These publications do not include the most recent dataset addition from
HVH, which will be described in a future publication.

\subsection{Patterns of Contralateral Involvement}\label{sec-data-strat}

These datasets allow us to investigate correlations between the
involvement of individual LNLs, or between risk factors and involvement.
In figure~\ref{fig-data-strat}, we have plotted the prevalence of each
contralateral LNL's involvement, stratified by the risk factors
T-category, number of ipsilaterally involved LNLs, and whether the tumor
extended over the mid-sagittal plane.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_data_strat.pdf}}

}

\caption{\label{fig-data-strat}Contralateral involvement stratified by
T-category (top left panel), the number of metastatic LNLs ipsilaterally
(top right panel), and whether the primary tumor extended over the
mid-sagittal line or was clearly lateralized (bottom left panel).
Additionally, in the bottom right, we show that even for lateralized
tumors, T-category and ipsilateral involvement are correlated with
contralateral metastasis by comparing the contralateral involvement
prevalence for selected scenarios that vary in their T-category and
ipsilateral involvement extent.}

\end{figure}%

The bottom left panel in figure~\ref{fig-data-strat} shows that patients
with a tumor crossing the mid-sagittal plane show contralateral
involvement vastly more often compared to patients with clearly
lateralized tumors. This makes intuitive sense, because the lymphatic
system in the head and neck region is typically symmetric and no major
lymph vessels cross the midline. Therefore, interstitial fluids from the
primary tumor -- which we assume to carry living malignant cells -- may
only reach the blind-ended lymphatic vessels in the contralateral neck
via short-ranged diffusion. Which in turn is only possible when the
primary tumor is close enough to the mid-sagittal line or crosses it.

The top left panel in figure~\ref{fig-data-strat} indicates that
T-category is correlated with contralateral involvement (as it is with
overal involvement). This is because T-category may be considered a
surrogate for the time between onset of disease and diagnosis. I.e., a
patient with a T4 tumor was -- on average -- diagnosed later than a
patient with a T1 tumor. Thus, the former did have more time to develop
metastases.

Similarly, ipsilateral involvement correlates with contralateral
metastasis (top right panel). The tumor of a patient with many
metastases in ipsilateral LNLs was probably able to spread for longer
(or faster) compared to a tumor in a patient with no nodal disease. The
degree of ipsilateral involvement, too, may therefore be considered a
surrogate for the duration of the disease. In addition, it has been
hypothesized that bulky nodal disease ipsilaterally may also redirect
lymph fluids to the contralateral side.

The three risk factors for contralateral involvement, midline extension,
T-category, and ipsilateral involvement are correlated. For example,
midline extension was present in 45.56\% of advanced T-category tumors
but only in 6.13\% of early T-category tumors. The higher prevalence of
contralateral metastases in advanced T-category tumors may thus be
partially explained by a higher rate of midline extension in these
tumors. However, the bottom right panel of figure~\ref{fig-data-strat}
indicates that advanced T-category and severe ipsilateral involvement do
represent additional risk factors. The figure considers only patients
with lateralized tumors not extending over the midline. For early
T-category and healthy ipsilateral levels I-V, only 1.85\% (2 out of 108
patients) shows involvement of contralateral level II. This increases to
6.73\% (14 out of 208 patients) if ipsilateral level II is involved, to
14.29\% (10 out of 70 patients) if ipsilateral levels II and III are
involved, and to 39.68\% (25 out of 63 patients) for advanced T-category
with ipsilateral levels II and III involved.

\subsection{Requirements for a Bilateral Model}\label{sec-requirements}

Based on the observations of the \hyperref[sec-data-strat]{previous
section}, any model to predict the risk for contralateral nodal
involvement, should be able to describe the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A tumor that extends over the mid-sagittal line should yield
  contralateral metastases with much higher probability.
\item
  More advanced T-category should lead to higher risk for nodal disease.
  One approach to achieve this via the expected time of diagnosis has
  already been developed in the form of a hidden Markov model
  \citep{ludwig_hidden_2021}.
\item
  The degree of ipsilateral involvement should give the model
  information on the time that may have passed between onset and
  diagnosis of the disease. This should come in addition to what can be
  inferred about the time from T-category alone.
\end{enumerate}

\section{Unilateral Model for Lymphatic
Progression}\label{sec-unilateral}

This paper builds on the previously developed unilateral model for
ipsilateral lymph node involvement presented in {[}cite 2024 graph
extension paper{]}. In this section we provide a brief recap of the
unilateral model to introduce the notation needed to extend the
framework to a bilateral model describing both ipsilateral and
contralateral lymph node involvement in section~\ref{sec-ext-to-contra}.
For further details in the ipsilateral model, the reader is referred to
the earlier publications \citep[cite 2024 graph extension
paper]{ludwig_hidden_2021}

We model a patient's state of involvement at an abstract time-step \(t\)
as a vector of hidden binary random variables:

\begin{equation}\phantomsection\label{eq-state-def}{
\mathbf{X}[t] = \begin{pmatrix} X_v[t] \end{pmatrix} \qquad v \in \left\{ 1, 2, \ldots, V \right\}
}\end{equation}

Here, \(V\) is the number of LNLs the model considers. The values a
LNL's hidden binary random variables may take on are \(X_v[t] = 0\)
(\texttt{False}), meaning the LNL \(v\) is healthy or free of metastatic
disease, or \(X_v[t] = 1\) (\texttt{True}), corresponding to some form
of tumor presence (i.e., occult or clinically detected). Since the state
vector \(\mathbf{X}[t]\) is \(V\)-dimensional and binary, there are
\(2^V\) distinct possible lymphatic involvement patterns, which we
enumerate from
\(\boldsymbol{\xi}_0 = \begin{pmatrix} 0 & 0 & \cdots & 0 \end{pmatrix}\)
to
\(\boldsymbol{\xi}_{2^V} = \begin{pmatrix} 1 & 1 & \cdots & 1 \end{pmatrix}\).
In addition, each LNL is associated with an observed binary random
variable \(Z_v\) that describes the clinical involvement of a LNL based
on imaging: \(Z_v = 0\) (\texttt{False}) indicates that the LNL \(v\) is
healthy based on clinical diagnosis, and \(Z_v = 1\) (\texttt{True})
indicates that suspicious lymph nodes were detected that are deemed
metastatic.

Any hidden Markov model is fully described by three quantities:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A starting state \(\mathbf{X}[t=0]\) at time \(t=0\) just before the
  patient's tumor formed. In our case, this is always the state where
  all LNLs are still healthy \(\boldsymbol{\xi}_0\).
\item
  The \emph{transition matrix}
  \begin{equation}\phantomsection\label{eq-trans-matrix}{
  \mathbf{A} = \left( A_{ij} \right) = \big( P \left( \mathbf{X}[t+1] = \boldsymbol{\xi}_j \mid \mathbf{X}[t] = \boldsymbol{\xi}_i \right) \big)
  }\end{equation} where the value at row \(i\) and column \(j\)
  represents the probability to transition from state
  \(\boldsymbol{\xi}_i\) to \(\boldsymbol{\xi}_j\) during the time-step
  from \(t\) to \(t+1\). Note that we prohibit self-healing, meaning
  that during a transition, no LNL may change their state from
  \(X_v[t]=1\) to \(X_v[t+1]=0\). Consequently, many elements of the
  transition matrix are zero.
\item
  Lastly, the \emph{observation matrix}
  \begin{equation}\phantomsection\label{eq-obs-matrix}{
  \mathbf{B} = \left( B_{ij} \right) = \big( P \left( \mathbf{Z} = \boldsymbol{\zeta}_j \mid \mathbf{X}[t_D] = \boldsymbol{\xi}_i \right) \big)
  }\end{equation} where in row \(i\) and at column \(j\) we find the
  probability to \emph{observe} a lymphatic involvement pattern
  \(\mathbf{Z} = \boldsymbol{\zeta}_j\), given that the true (but
  hidden) state of involvement at the time of diagnosis \(t_D\) is
  \(\mathbf{X}[t_D] = \boldsymbol{\xi}_i\).
\end{enumerate}

The transition matrix \(\mathbf{A}\) is parametrized using a directed
acyclic graph (DAG) as an abstract representation of the underlying
lymphatic network. Directed arcs from the primary tumor to a LNL are
associated with a probability \(b_v\) for direct spread to LNL \(v\)
during one time step. Directed arcs from a LNL \(v\) to a LNL \(r\) are
associated with a probability \(t_{vr}\) for the tumor to progress from
one LNL to the next. In this paper, we build on the DAG shown in
figure~\ref{fig-full-graph} which was obtained by maximizing the model
evidence as described in {[}cite 2024 graph extension paper{]}.

\begin{figure}

\centering{

\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/static/full-graph.pdf}

}

\caption{\label{fig-full-graph}Directed acyclic graph (DAG) representing
the abstract lymphatic network in the head and neck region. Blue nodes
are the LNLs' hidden random variables, the red node represents the
tumor, and the orange square nodes depict the binary observed variables.
Red and blue arcs symbolize the probability of lymphatic spread along
that edge during one time-step. The orange arcs represent the
sensitivity and specificity of the observational modality (e.g.~CT, MRI,
pathology, \ldots).}

\end{figure}%

Using the introduced quantities, we can evolve the distribution of all
possible hidden states from \(\mathbf{X}[t=0] = \boldsymbol{\xi}_0\)
step by step, by successively multiplying this vector with the
transition matrix \(\mathbf{A}\). For later use, we define at this point
the matrices of the dsitributions over all hidden states, given all
time-steps \(\boldsymbol{\Lambda}\):

\begin{equation}\phantomsection\label{eq-lambda-matrix}{
\boldsymbol{\Lambda} = P \left( \mathbf{X} \mid \mathbf{t} \right) = \begin{pmatrix}
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^0 \\
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^1 \\
\vdots \\
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^{t_\text{max}} \\
\end{pmatrix}
}\end{equation}

Where the \(k\)-th row in this matrix corresponds to the probability
distribution over hidden states after \(t=k-1\) time-steps.

At the time of the diagnosis \(t_D\), we multiply the result with the
observation matrix \(\mathbf{B}\). We may then look up the likelihood of
a patient presenting with the diagnosis
\(\mathbf{Z}=\boldsymbol{\zeta}_i\) in the \(i\)-th entry of the final
result. However, the remaining issue is that the value of \(t_D\) is
unknown, i.e.~over how many time-steps the HMM should be evolved. We
solve this problem by marginalizing over the time of diagnosis.
Different distributions over the diagnosis times can then be choosen
based on T-category. For instance, the mean of the time-prior to
marginalize over the diagnosis time for early T-category patients
\(P\left( t_D \mid \text{early} \right)\) may be shifted towards earlier
times than the one for advanced T-category patients
\(P\left( t_D \mid \text{early} \right)\). This gives us for example

\[
P\left( \mathbf{X} \mid \text{T}x = \text{early} \right) = \sum_{t=0}^{t_\text{max}} P \left( \mathbf{X} \mid t \right) \cdot P(t \mid \text{early})
\]

In this work, we use binomial distributions
\(\mathfrak{B} \left( t_D, p_{\text{T}x} \right)\) as time-priors which
have one free parameter \(p_{\text{T}x}\) for each group of patients we
differentiate based on T-category. Also, we fix \(t_\text{max} = 10\),
which means that the expected number of time-steps from the onset of a
patient's disease to their diagnosis is
\(\mathbb{E}\left[ t_D \right] = 10 \cdot p_{\text{T}x}\).

\subsection{Likelihood Function of the Unilateral
Model}\label{likelihood-function-of-the-unilateral-model}

With the formalism introduced above, we can write the likelihood
function for a patient to present with a diagnosis consisting of an
observed state and a T-category
\(d = \left( \boldsymbol{\zeta}_i, \text{T}x \right)\) as follows:

\begin{equation}\phantomsection\label{eq-single-patient-llh}{
\ell = P \left( \mathbf{Z} = \boldsymbol{\zeta}_i \mid \text{T}x \right) = \sum_{t=0}^{t_\text{max}} \left[ \boldsymbol{\xi}_0 \cdot \mathbf{A}^t \cdot \mathbf{B} \right]_i \cdot P \left( t \mid \text{T}x \right)
}\end{equation}

Above, the quantity inside \(\left[ \ldots \right]_i\) denotes the
\(i\)-th component of the vector that is the result of the vector and
matrix multiplications in the square brackets. Note that it is also
possible to account for missing involvement information: If a diagnosis
(like fine needle aspiration (FNA)) is only available for a subset of
all LNLs, we can sum over all those possible complete observed states
\(\boldsymbol{\zeta}_j\) that match the provided diagnosis.

The single-patient likelihood \(\ell\) in
equation~\ref{eq-single-patient-llh} depends on the spread parameters
shown in figure~\ref{fig-full-graph} via the transition matrix
\(\mathbf{A}\) and on the binomial parameters \(p_{\text{T}x}\) via
time-priors. In this work, we will only differentiate between ``early''
(T1 \& T2) and ``advanced'' (T3 \& T4) T-categories. Therefore, the
parameter space of the unilateral model is:

\begin{equation}\phantomsection\label{eq-param-space}{
\boldsymbol{\theta} = \left( \left\{ b_v \right\}, \left\{ t_{vr} \right\}, p_\text{early}, p_\text{adv.} \right) \quad \text{with} \quad \genfrac{}{}{0pt}{2}{v\leq V}{r\in\operatorname{pa}(v)}
}\end{equation}

And it is our goal to infer the values of these parameters for a given
dataset \(\mathcal{D} = \left( d_1, d_2, \ldots, d_N \right)\) of OPSCC
patients. The likelihood of these \(N\) diagnoses is simply the product
of their individual likelihoods as defined in
equation~\ref{eq-single-patient-llh}. For numerical reasons, we
typically compute the data likelihood in log space:

\begin{equation}\phantomsection\label{eq-log-likelihood}{
\log \mathcal{L} \left( \mathcal{D} \mid \boldsymbol{\theta} \right) = \sum_{i=1}^N \log \ell_i
}\end{equation}

The methodology we use to infer the model's parameters is detailed in
section~\ref{sec-sampling}.

\section{Extension to a Bilateral Model}\label{sec-ext-to-contra}

A naive approach to model the contralateral lymphatic spread would be to
simply employ two independent unilateral models as introduced in
section~\ref{sec-unilateral}. During training, one could enforce that
some parameters are shared between these two models, e.g.~the
parameterization of the distributions over diagnose times or the spread
among the LNLs (\$t\_\{vr\}). However, this approach lacks a way to
describe the correlation between ipsi- and contralateral involvement
discussed in section \{\#sec-data-strat\}. Two independent unilateral
models would not describe the observation that contralateral involvement
becomes more likely with more severe ipsilateral involvement.

Thus, we extend the formalism in section~\ref{sec-unilateral} in such a
way that the model's ipsi- and contralateral side evolve synchronously
over time. To achieve that, we start by writing down the posterior
distribution of involvement an analogy to
equation~\ref{eq-uni-bayes-law}, which is now a joint probability of an
involvement \(\mathbf{X}^\text{i}\) ipsilaterally \emph{and} an
involvement \(\mathbf{X}^\text{c}\) contralaterally, given a diagnosis
of the ipsilateral LNLs \(\mathbf{Z}^\text{i}\) and of the contralateral
ones \(\mathbf{Z}^\text{c}\):

\begin{equation}\phantomsection\label{eq-bilateral-bayes}{
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right) = \frac{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right)}{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right)}
}\end{equation}

For the sake of brevity, we omit the dependency on the parameters and
the T-category here.

The probability of the diagnoses given a hidden state factorises:
\(P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) = P \left( \mathbf{Z}^\text{i} \mid \mathbf{X}^\text{i} \right) \cdot P \left( \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{c} \right)\),
and the two factors are described through observation matrices
\(\mathbf{B}^\text{i}\) and \(\mathbf{B}^\text{c}\).

The term representing the model's prior probability of hidden
involvement does not factorize. However, we assume that there is no
direct lymph drainage from an ipsilateral LNL to a contralateral LNL. No
major lymph vessels cross the mid-sagittal plane. In the graphical
model, this means that we assume no directed arcs between ipsilateral
and contralateral LNLs, i.e.~tumor spread to the contralateral side is
due to the primary tumor alone. We can thus write the joint probability
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right)\) as a
factorising sum:

\begin{equation}\phantomsection\label{eq-bilateral-marginal}{
\begin{aligned}
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) &= \sum_{t=0}^{t_\text{max}} P(t) \cdot P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid t \right) \\
&= \sum_{t=0}^{t_\text{max}} P(t) \cdot P \left( \mathbf{X}^\text{i} \mid t \right) \cdot P \left( \mathbf{X}^\text{c} \mid t \right)
\end{aligned}
}\end{equation}

This assumption makes intuitive sense: The state of the ipsilateral and
contralateral sides of the lymphatic network evolves independently over
time as there are no major lymph vessels crossing the midline. However,
both sides are coupled via time. Qualitatively speaking, according to
the model in \{\#eq-bilateral-marginal\}, a joint state with severe
contralateral involvement and limited ipsilateral involvement would be
unlikely, because severe contralateral involvement becomes likely only
at later time steps when limited ipsilateral involvement becomes
unlikely.

Using equation~\ref{eq-bilateral-marginal} along with
equation~\ref{eq-lambda-matrix}, we can write the above distribution
algebraically as a product:

\begin{equation}\phantomsection\label{eq-bilateral-marginal-algebra}{
P \left( \mathbf{X}^\text{i} = \boldsymbol{\xi}_n, \mathbf{X}^\text{c} = \boldsymbol{\xi}_m \right) = \left[ \boldsymbol{\Lambda}^\intercal_\text{i} \cdot \operatorname{diag} P(\mathbf{t}) \cdot \boldsymbol{\Lambda}_\text{c} \right]_{n,m}
}\end{equation}

\subsection{Parameter Symmetries}\label{sec-params-symmetry}

In general, the matrices \(\boldsymbol{\Lambda}_\text{i}\) and
\(\boldsymbol{\Lambda}_\text{c}\) could be parameterized using a
disjoint set of parameters, i.e., the ipsi- and contralateral spread
rates are entirely different. However, using three sensible assumptions,
we can reduce the parameter space by sharing some parameters between the
two sides:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We assume that both ipsilateral and contralateral spread is described
  through the same graph shown in \{\#fig-full-graph\}.
\item
  We assume the spread \emph{among} the LNLs to be same on both sides.
  It is reasonable to assume the lymphatic system is symmetric. Thus,
  the spread rates from one LNL to the other should be symmetric, too.
  Formally, this means\\
  \begin{equation}\phantomsection\label{eq-symmetries}{
  \begin{aligned}
  b_v^\text{c} &\neq b_v^\text{i} \\
  t_{rv}^\text{c} &= t_{rv}^\text{i}
  \end{aligned}
  }\end{equation} for all \(v \leq V\) and
  \(r \in \operatorname{pa}(v)\).
\item
  The tumor's spread to the contralateral side in case of an extension
  over the midline is larger than if it was clearly lateralized, but
  smaller than its spread to the ipsilateral side. This assumption stems
  from a simple thought experiment: Consider moving the tumor from a
  clearly lateralized position accross the mid-sagittal plane to the
  same position, but on the contralateral side. In the beginning we
  would have \(b_v^\text{c} < b_v^\text{i}\), while in the end, the
  situation is reversed. If a tumor extends over the mid-sagittal line,
  its contralateral spread rate can be expected to be in between these
  two extremes. We encode this in a \emph{mixing parameter}
  \(\alpha \in [0,1]\) that captures a ``degree of asymmetry'':\\
  \begin{equation}\phantomsection\label{eq-mixing}{
  b_v^{\text{c},\epsilon=\texttt{True}} = \alpha \cdot b_v^\text{i} + (1 - \alpha) \cdot b_v^{\text{c},\epsilon=\texttt{False}}
  }\end{equation} This means the model now uses three different sets of
  parameters to describe the spread from the tumor to the LNLs:
  \(b^\text{i}_v\) for the spread to the ipsilateral LNLs,
  \(b_v^{\text{c},\epsilon=\texttt{False}}\) for the spread to the
  contralateral LNLs as long as the tumor is clearly lateralized, and
  finally \(b_v^{\text{c},\epsilon=\texttt{True}}\) when it crosses the
  midline. Note, however, that these three sets of spread rates only
  account for \(2 \cdot 2^V + 1\) parameters, since they are coupled via
  the mixing parameter \(\alpha\).
\end{enumerate}

Our parameter space has now expanded to

\begin{equation}\phantomsection\label{eq-bi-param-space}{
\boldsymbol{\theta} = \left( \left\{ b_v^\text{i} \right\}, \left\{ b_v^\text{c} \right\}, \alpha, \left\{ t_{vr} \right\}, p_\text{early}, p_\text{adv.} \right) \quad \text{with} \quad \genfrac{}{}{0pt}{2}{v\leq V}{r\in\operatorname{pa}(v)}
}\end{equation}

which is less than double in size, compared to the unilateral model.
From these parameters, we now build three different transition matrices:
The unchanged matrix \(\mathbf{A}_\text{i}\) for the ipsilateral side,
one \(\mathbf{A}_\text{c}^{\epsilon=\texttt{False}}\) for the
contralateral side that covers the progression as long as the tumor is
lateralized, and \(\mathbf{A}_\text{c}^{\epsilon=\texttt{True}}\) in
case of a tumor that has crossed the mid-sagittal line.

\subsection{Modelling Midline Extension}\label{sec-midline}

It can be assumed that most tumors that cross the midline at the time of
diagnose have started as lateralized tumors and grew over the midline at
a later time step. Hence, the transition matrix
\(\mathbf{A}_\text{c}^{\epsilon=\texttt{True}}\) only applies for a
subset of the time steps. Therefore, we also model the tumor's extension
over the mid-sagittal line as a binary random variable. A tumor starts
lateralized and at every time-step there is a finite probability
\(p_\epsilon\) that the tumor grows over the midsaggital plane. The
overall probabilities to find a patient with a clearly lateralized tumor
or one that extends over the mid-sagittal line after \(t\) time-steps
are then respectively given by

\[
\begin{aligned}
P(\epsilon = \texttt{False} \mid t) &= (1 - p_\epsilon)^t \\
P(\epsilon = \texttt{True} \mid t) &= 1 - P(\epsilon = \texttt{False} \mid t)
\end{aligned}
\]

Using this, it is straightforward to write down the matrix of state
distributions for all time-steps, as in equation~\ref{eq-lambda-matrix}
covering the contralateral hidden state evolution:

\[
\boldsymbol{\Lambda}_\text{c}^{\epsilon=\texttt{False}} =
\left(
\begin{array}{r}
\boldsymbol{\pi}^\intercal \cdot \left( \mathbf{A}_\text{c}^{\epsilon=\texttt{False}} \right)^{0\phantom{t_\text{max}}} \\
(1-p_\epsilon) \cdot \boldsymbol{\pi}^\intercal \cdot \left( \mathbf{A}_\text{c}^{\epsilon=\texttt{False}} \right)^{1\phantom{t_\text{max}}} \\
\hfill \vdots \hfill \\
(1-p_\epsilon)^{t_\text{max}} \cdot \boldsymbol{\pi}^\intercal \cdot \left( \mathbf{A}_\text{c}^{\epsilon=\texttt{False}} \right)^{t_\text{max}\phantom{0}} \\
\end{array}
\right)
\]

where we used the transition matrix
\(\mathbf{A}_\text{c}^{\epsilon=\texttt{False}}\) that depends on the
base spread parameters \(b_v^{\text{c},\epsilon=\texttt{False}}\).

The case when midline extension is eventually present is more
complicated: We already marginalize over the exact time-step when the
tumor grows over the mid-sagittal line. But whenever that happens, we
also need to change the contralateral transition matrix to use the
increased spread rates \(b_v^{\text{c}, \epsilon=\texttt{True}}\) from
the tumor to the contralateral LNLs, given by the linear mixing in
equation~\ref{eq-mixing}. We can achieve the correct marginalization by
iteratively building the joint distribution
\(P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid t \right)\).
We start from \(t=0\), where we know that all LNLs are healthy (the
contralateral neck is in the state
\(\mathbf{X}_\text{c}=\boldsymbol{\xi}_0\)) and the tumor is lateralized
(\(\epsilon=\texttt{False}\)):

\[
P \left( \mathbf{X}^\text{c} = \boldsymbol{\xi}_0, \epsilon=\texttt{False} \mid t=0 \right) = 1
\]

whereas all other states have zero probability. Then we consider an
arbitrary later time-step \(t=\tau+1\). There are two possible scenarios
we need to marginalize over:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The tumor was still lateralized at \(t=\tau\) and just grew over the
  midline. Under this scenario, the probability for having a midline
  extension at \(t=\tau+1\) is given by \(p_\epsilon\). We use this to
  weight the contralateral state distribution that was so far evolved
  without increased contralateral spread.
\item
  The mid-sagittal line was already crossed by the tumor. In that case,
  the probability is 1 to remain in that lateralization state. Thus, to
  account for this case we simply add the distribution
  \(P\left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid \tau \right)\)
  from one time-step earlier, arriving at a recursion relation:
\end{enumerate}

\[
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid \tau + 1 \right) = \big[ p_\epsilon P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{False} \mid \tau \right) + P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid \tau \right) \big]^\top \cdot \mathbf{A}_\text{c}^{\epsilon=\texttt{True}}
\]

We can collect the iteratively computed distributions for the midline
extension case to define the matrix over the states given all
time-steps, in analogy to equation~\ref{eq-lambda-matrix}:

\[
\boldsymbol{\Lambda}_\text{c}^{\epsilon=\texttt{True}} = \begin{pmatrix}
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid 0 \right) \\
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid 1 \right) \\
\vdots \\
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid t_\text{max} \right) \\
\end{pmatrix}
\]

Using this, we can again write the joint of ipsi- and contralateral
involvement - now also for the case of mid-sagittal extension -
algebraically as before in equation~\ref{eq-bilateral-marginal-algebra}:

\[
P \left( \mathbf{X}^\text{i} = \boldsymbol{\xi}_n, \mathbf{X}^\text{c} = \boldsymbol{\xi}_m, \epsilon \right) = \left[ \boldsymbol{\Lambda}^\intercal_\text{i} \cdot \operatorname{diag} P(\mathbf{t}) \cdot \boldsymbol{\Lambda}_\text{c}^\epsilon \right]_{n,m}
\]

We can use this term to compute the likelihood of all patients with and
without midline extension separately. And if for some patients the
information of tumor lateralization is not available, we can simply sum
the above term once for \(\epsilon=\texttt{False}\) and once for
\(\epsilon=\texttt{True}\) to marginalize over the unknown variable.

The final parameter space of our extended model has now reached this
size:

\begin{equation}\phantomsection\label{eq-ext-param-space}{
\boldsymbol{\theta} = \left( \left\{ b_v^\text{i} \right\}, \left\{ b_v^\text{c} \right\}, \alpha, \left\{ t_{vr} \right\}, p_\text{early}, p_\text{adv.}, p_\epsilon \right) \quad \text{with} \quad \genfrac{}{}{0pt}{2}{v\leq V}{r\in\operatorname{pa}(v)}
}\end{equation}

\subsection{Model Prediction in the Bayesian
Context}\label{model-prediction-in-the-bayesian-context}

Our stated goal is to compute the risk for a patient's true ipsi- and
contralateral nodal involvement states \(\mathbf{X}^\text{i}\) and
\(\mathbf{X}^\text{c}\), \emph{given} their individual diagnosis
\(d = \left( \boldsymbol{\zeta}^\text{i}_k, \boldsymbol{\zeta}^\text{c}_\ell, \epsilon, \text{T}x \right)\).
Here, this diagnosis consists of the observed ipsi- and contralateral
nodal involvements, the patient's midline extension \(\epsilon\), and
their tumor's T-category \(\text{T}x\). Using Bayes' law, we can write
this risk as:

\begin{equation}\phantomsection\label{eq-uni-bayes-law}{
P \big( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid d, \boldsymbol{\hat{\theta}} \big)
= \frac{P \left( \boldsymbol{\zeta}^\text{i}_k \mid \mathbf{X}^\text{i} \right) P \left( \boldsymbol{\zeta}^\text{c}_\ell \mid \mathbf{X}^\text{c} \right) P \big( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \mid \boldsymbol{\hat{\theta}}, \text{T}x \big)}
{\sum_{i=0}^{2^V} \sum_{j=0}^{2^V} \mathcal{C}_{ij}}
}\end{equation}

with the normalization constants

\[
\mathcal{C}_{ij} = P \left( \boldsymbol{\zeta}^\text{i}_k \mid \mathbf{X}^\text{i}=\boldsymbol{\xi}^\text{i}_i \right) P \big( \boldsymbol{\zeta}^\text{c}_\ell \mid \mathbf{X}^\text{c}=\boldsymbol{\xi}^\text{c}_j \big) P \big( \mathbf{X}^\text{i}=\boldsymbol{\xi}^\text{i}_i, \mathbf{X}^\text{c}=\boldsymbol{\xi}^\text{c}_j, \epsilon \mid \boldsymbol{\hat{\theta}}, \text{T}x \big)
\]

The terms
\(P \left( \boldsymbol{\zeta}^\text{i}_k \mid \mathbf{X}^\text{i} \right)\)
and
\(P \left( \boldsymbol{\zeta}^\text{c}_\ell \mid \mathbf{X}^\text{c} \right)\)
are defined solely by sensitivity and specificity of the diagnostic
modality. Terms like these already appeared in the definition of the
observation matrx in equation~\ref{eq-obs-matrix}. The \emph{prior}
\(P \big( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \mid \boldsymbol{\hat{\theta}}, \text{T}x \big)\)
in the above equation is the crucial term that is supplied by a trained
model and its parameters \(\boldsymbol{\hat{\theta}}\).

It is possible to compute this \emph{posterior} probability of true
involvement not only for one fully defined state
\((\mathbf{X}^\text{i}, \mathbf{X}^\text{c})\), but also for
e.g.~individual LNLs: For example, the risk for involvement in the
contralateral level IV would be a marginalization over all ipsilateral
states \(\boldsymbol{\xi}^\text{i}_i\) and all contralateral states
\(\boldsymbol{\xi}^\text{c}_j\) where \(\xi^\text{c}_{j4}=1\). Formally:

\begin{equation}\phantomsection\label{eq-marg-over-posterior}{
P \big( \text{cIV} \mid \mathbf{Z}^\text{i}=\boldsymbol{\zeta}^\text{i}_k, \mathbf{Z}^\text{c}=\boldsymbol{\zeta}^\text{c}_\ell, \boldsymbol{\hat{\theta}}, \text{T}x \big) = \sum_k \sum_{\ell \, : \, \xi_{\ell 4}=1} P \big( \mathbf{X}^\text{i} = \boldsymbol{\xi}^\text{i}_k, \mathbf{X}^\text{c} = \boldsymbol{\xi}^\text{c}_\ell \mid \boldsymbol{\zeta}^\text{i}_k, \boldsymbol{\zeta}^\text{c}_\ell, \epsilon, \boldsymbol{\hat{\theta}}, \text{T}x \big)
}\end{equation}

\section{Computational methods}\label{sec-methods}

In this section, we detail how the experiments were performed. Every
figure, table, and result is fully reproducible via the GitHub
repository
\href{https://github.com/rmnldwg/bilateral-paper}{\texttt{rmnldwg/bilateral-paper}}.
It also contains the raw manuscript and instructions on how to recreate
all figures, tables, and the final document.

\subsection{Involvement Data Consensus}\label{sec-methods-consensus}

It is possible to provide our model with multiple different diagnostic
modalities, each being characterized by different pairs of sensitivity
and specificity. However, we instead chose to combine them into a single
``consensus'' diagnosis before parameter inference, as introduced in
section~\ref{sec-data-consensus}. We opted for this because the
literature values of sensitivity and specificity
\citep{de_bondt_detection_2007, kyzas_18f-fluorodeoxyglucose_2008} of
imaging modalities like MRI and CT do not plausibly match some of our
observations: In the USZ cohort, 78\% of OPSCC patients where diagnosed
with ipsilateral LNL II involvement via diagnostic imaging. This is
virtually impossible with sensitivities around 80\% and specificities
lower than 100\%.

\subsection{MCMC Sampling}\label{sec-sampling}

For parameter inference, we used the Python package
\href{https://emcee.readthedocs.io/en/stable/}{\texttt{emcee}}
\citep{foreman-mackey_emcee_2013}. It implements efficient MCMC sampling
algorithms that employ multiple parallel samplers for affine invariance
and better performance on multi-core CPUs. The sample proposal
algirothms used by us are based on differential evolution moves
\citep{ter_braak_differential_2008, nelson_run_2013}. The
\href{https://emcee.readthedocs.io/en/stable/}{\texttt{emcee}} library
was provided with the likelihood implemented by our
\href{https://lymph-model.readthedocs.io/en/stable/}{\texttt{lymph-model}}
Python package.

For each dimension in the parameter space of the model, we initialized
12 of these parallel samplers, called ``walkers'', with random values in
the unit cube. Every time all of these walkers advanced 50 steps, the
autocorrelation time of the chains was estimated. For short chains, this
estimate is not trustworthy, but stabilizes for longer chains. We
therefore considered a sampling to be converged when two criteria were
met:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The change in the autocorrelation time was less then 5.0e-2.
\item
  The estimate of the autocorrelation dropped below \(n\) / 50 where
  \(n\) is the length of the chain up to that point.
\end{enumerate}

All samples up to this convergence - called the \emph{burn-in phase} -
were discarded. We only kept another 10 samples after that, which were
spaced 10 steps apart.

First, in figure~\ref{fig-model-burnin-history}, we verify the sampling
converged successfully by inspecting two monitoring quantities: The
autocorrelation time of the MCMC chain and the acceptance fractions of
the parallel walkers.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_burnin_history.pdf}}

}

\caption{\label{fig-model-burnin-history}Monitoring quantities during
the burn-in phase of the parameter sampling. Left: The autocorrelation
time of the sampling chain estimated at different sampling steps. We
consider the chain converged when the estimate of the autocorrelation
time is stable and drops below the trust threshold of \(n/50\) where
\(n\) is the number of steps. Right: Fraction of accepted MCMC proposals
averaged over all parallel walkers. Values around 30\% indicate good
mixing of the walkers.}

\end{figure}%

\subsection{Computing the Observed and Predicted Prevalence of
Involvement Patterns}\label{sec-prevalence}

We want to assess the model's capability to approximate the distribution
of lymphatic involvement patterns seen in the data. To that end, we
compare the prevalence of some invovlement patterns under selected
scenarios with the model's prediction for how often these involvements
it expects to see, given these scenarios.

In this context, a ``scenario'' includes the patient's T-category
\(\text{T}x\) and whether the patient's tumor extended over the
mid-sagittal line, i.e.~\(\epsilon=\texttt{True}\) or
\(\epsilon=\texttt{False}\).

An involvement pattern specifies for each ipsi- and contralateral LNL
whether it is ``healthy'', ``involved'', or ``masked''. If it is
``masked'', we essentially state that we are not interested in the
involvement of that LNL and the prevalence will be marginalized over
this LNL's involvement.

For example, we may be interested in the prevalence of contralateral LNL
II involvement (i.e., contra LNL II ``involved'' and all other LNLs
``masked'') under the scenario of early T-category (T0-T2) and no
midline extension (\(\epsilon=\texttt{False}\)). To compute this
prevalence in the data, we select all patients of this scenario (in our
data, this amounts to 379 patients). Of those, 27 were found to harbor
metastases in their contralateral LNL II. Therefore, the prevalence is
7.12\%.

When displaying this data prevalence, we often choose to draw a
\emph{beta posterior} over the ``true'' prevalence, hinting at the fact
that our data merely represents a limited sample. The beta posterior
follows from a uniform beta distribution as prior and a binomial
likelihood for the number of patients with the involvement of interest,
given the parameer for the ``true'' prevalence. The resulting
distribution has its maximum at the observed prevalence, but in addition
gives a visual intuition for the variance of of the observed quantity.
I.e., when we observe 3 out of 10 events, the beta posterior is much
wider than if we observe 300 out of 1000 for he same prevalence. It also
allows us to check not only if the model is accurate, but also whether
it reflects the uncertainty contained in the data.

Predicting the prevalence using our model amounts to computing the
following probability:

\[
P \left( \text{II}^\text{c} \mid \epsilon=\texttt{False}, \text{T}x=\text{early} \right) = \frac{P \left( \text{II}^\text{c}, \epsilon=\texttt{False} \mid \text{T}x=\text{early} \right)}{P \left( \epsilon=\texttt{False} \mid \text{T}x=\text{early} \right)}
\]

In the enumator, we marginalize over all ipsi- and contralateral LNLs'
involvements, except for LNL II contralaterally. This is similar to the
marginalization in equation~\ref{eq-marg-over-posterior}, although we
are summing over different quantities. In the denominator, we can simply
insert the joint distribution over midline extension and diagnose time
\(P \left( \epsilon, t \right)\) marginalized over \(t\) using the early
T-category's time-prior.

Since we compare it to the data, which does not report true but only
observed involvement -- although pathologically investigated LNLs may be
as close as possible to the ground truth -- we do not consider
posteriors of the form \(P \left( \mathbf{X} \mid \mathbf{Z} \right)\)
here. Instead, we compute probabilities of observed involvement
\(P \left( \mathbf{Z} \right)\), as in the likelihood
equation~\ref{eq-single-patient-llh}.

When plotted, we usually display histograms over the model's
predictions. Each of their values was computed from a different
parameter set drawn during MCMC sampling, effectively giving us a
distribution over the prevalences. Ideally, the histograms approximate
the location and width of the Beta posteriors when attempting to
describe the data they were trained on.

Note that we decided to omit the y-axis ticks and labels in these
figures over prevalences and risks. The y-axis in these plots measures
the probability density and its numerical values are not intuitively
interpretable. Instead, we occasionally use the freed space to label
e.g.~rows of subplots.

\section{Results: Model evaluation}\label{sec-results}

In table~\ref{tbl-midline-params}, we tabulate the mean and standard
deviation of the sampled parameters for the full midline model.
Considering the parameters of the bilateral model that describe the
ipsilateral spread (\(b^i_v\), \(t_{vr}\) and \(p_\text{adv.}\)), we
note that the bilateral model mostly reproduces the parameter values
reported in the earlier publication \citep{ludwig_modelling_2023} for
the unilateral model. The small differences may be mostly due to the
differences in the training dataset. Therefore, the analysis of the
model's capability to adequately describe ipsilateral lymph node
involvement is not repeated here and we focus instead on the analysis of
contralateral involvement. We can further note that the contralateral
spread parameters \(b^c_v\) are small compared to the ipsilateral
parameters \(b^i_v\), reflecting the low prevalence of contralateral
lymph node involvement for lateralized tumors.

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\begin{longtable}[]{@{}lll@{}}

\caption{\label{tbl-midline-params}Mean sampled parameter estimates of
the midline model and the respective standard deviation. The parameters
set to fixed values are the maximum number of time steps \(t_{max}=10\)
and the time prior parameter for early T-category patients
\(p_{early}=0.3\).}

\tabularnewline

\caption{}\label{T_7c6a5}\tabularnewline
\toprule\noalign{}
Parameter & Mean & Std. Dev. \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Parameter & Mean & Std. Dev. \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Mid. ext. probability & 8.16\% & ± 0.48\% \\
ipsi: T ➜ I & 2.80\% & ± 0.26\% \\
ipsi: T ➜ II & 34.89\% & ± 1.40\% \\
ipsi: T ➜ III & 5.45\% & ± 0.66\% \\
ipsi: T ➜ IV & 0.94\% & ± 0.18\% \\
ipsi: T ➜ V & 1.83\% & ± 0.22\% \\
ipsi: T ➜ VII & 2.32\% & ± 0.26\% \\
contra: T ➜ I & 0.29\% & ± 0.09\% \\
contra: T ➜ II & 2.46\% & ± 0.29\% \\
contra: T ➜ III & 0.14\% & ± 0.07\% \\
contra: T ➜ IV & 0.19\% & ± 0.08\% \\
contra: T ➜ V & 0.05\% & ± 0.04\% \\
contra: T ➜ VII & 0.50\% & ± 0.17\% \\
Mixing ⍺ & 33.87\% & ± 4.32\% \\
I ➜ II & 62.50\% & ± 16.69\% \\
II ➜ III & 14.23\% & ± 1.64\% \\
III ➜ IV & 15.86\% & ± 1.93\% \\
IV ➜ V & 14.58\% & ± 3.76\% \\
late T-cat. binom. prob. & 44.98\% & ± 1.99\% \\

\end{longtable}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\subsection{Illustration of the model}\label{illustration-of-the-model}

In this subsection we illustrate parts of the mathematical framework
defined in the previous section. In figure~\ref{fig-model-midext-evo}
(top panel), we visualize the prior distribution over diagnose times
\(P(t)\). By parameter choice, early T-category tumors are diagnosed on
average after 3 time steps. Advanced T-category tumors are diagnosed on
average after 4.5 time steps, because the parameter \(p_\text{adv.}\)
was learned to be 44.98\% for advanced T-category tumors. The
probability per time step for the tumor to grow over the midline
(\(p_\epsilon\)) was on average 8.16\%. Using this, one can compute the
probability of midline extension \(P(\epsilon \mid t)\) at a given time
step \(t\) (red line in figure~\ref{fig-model-midext-evo}). The bottom
panel in this figure shows the joint probability \(P(\epsilon, t)\) to
be diagnosed at time step \(t\) with a given state of midline extension
and T-category.

\begin{figure}

\centering{

\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/figures/fig_model_midext_evo.pdf}

}

\caption{\label{fig-model-midext-evo}The top panel shows the prior
probability to get diagnosed at time-step \(t\) for early and late
T-category tumors as bars. Also in the top panel, we plot the
conditional probability of the tumor's midline extension
(\(\epsilon=\texttt{True}\)), given the time-step \(t\) as a line plot.
In the bottom panel, we show the joint probability of getting diagnosed
in time-step \(t\) \emph{and} having a tumor that crosses the midline.}

\end{figure}%

The model introduced in the previous section defines a model of the
joint probability distribution over midline extension and ipsi- and
contralateral lymph node involvement,
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right)\).
How this quantity is computed is illustrated in
figure~\ref{fig-model-algebra-equation}, visually representing
equation~\ref{eq-bilateral-marginal-algebra}. To keep it manageable to
interpret, this example only considers the LNLs II, III, and IV ipsi-
and contralaterally, resulting in \(2^3 = 16\) distinct states per side
and thus \(2 \times 8 \times 8 = 128\) total states. LNLs I, V, and VII
and the associated spread parameters have been removed from the model
while the remaining parameters are set to the mean values in
table~\ref{tbl-midline-params}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_algebra_equation.pdf}}

}

\caption{\label{fig-model-algebra-equation}Visual representation of
equation~\ref{eq-bilateral-marginal-algebra} for the case of midline
extention. The left and right matrices represent the evolution of the
possible hidden states for the ipsi- and contralateral neck
respectively. In particular, the right matrices describe the evolution
of the contralateral hidden states once for the case of no midline
extension (top) and once for the case where the tumor does extend over
the mid-sagittal line (bottom). In the center, the time-prior for late
T-category is plotted as a diagonal matrix. The result of this matrix
computation is the joint distribution
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right)\),
visualized in figure~\ref{fig-model-state-dist}}

\end{figure}%

The matrix on the left side of figure~\ref{fig-model-algebra-equation}
shows the time evolution of the probability distribution over the
ipsilateral involvement state, starting from the healthy state
\([0,0,0]\) with probability one. The two matrices on the right side
show the time evolution of the probability distribution over the
contralateral involvement state, where we need to distinguish the two
cases with and without midline extension. At time step \(t=0\), the
contralateral neck starts from the healthy state \([0,0,0]\) without
midline extension. The matrix in the center shows the time prior for
late T-category. The result of the matrix multiplication is the joint
probability distribution over midline extension and ipsi- and
contralateral lymph node involvement,
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon=\texttt{True} \right)\)
shown in figure figure~\ref{fig-model-state-dist}. The joint
distribution is shown as two separate heatmaps for the two midline
extension states. The most likely state is a lateralized tumor with
involvement of ipsilateral level II and no contralateral involvement,
which has a probability of approximately 25\%. The second most probable
state is a lateralized tumor with ipsilateral involvement of levels II
and III and no contralateral involvement. The most likely state with
contralateral involvement are tumors with midline extension with
involvement of contralateral level II and ipsilateral level II and III.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_state_dist.pdf}}

}

\caption{\label{fig-model-state-dist}The full distribution
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon=\text{False} \right)\)
over the ipsi- and contralateral states, as well as midline extension as
computed by a model considering the LNLs II, III, and IV for a late
T-category tumor. Since this is a three-dimensional object with one
dimension being the binary random variable representing midline
extension \(\epsilon\), we show this distribution as separate
two-dimensional heatmaps. These two matrices are the result of
equation~\ref{eq-bilateral-marginal-algebra}, which we visualized as
well in figure~\ref{fig-model-algebra-equation}.}

\end{figure}%

\subsection{Prevalence predictions for contralateral
involvement}\label{prevalence-predictions-for-contralateral-involvement}

The bilateral model was designed to fulfil the requirements laid out in
section~\ref{sec-requirements}. Now, we investigate to what extent the
model can quantitatively describe the patterns of lymph node involvement
observed in the dataset. To that end, we compare the model's predictions
for contralateral involvement against observations in the data. This is
done given scenarios that differ in T-category and/or midline extension
and/or ipsilateral involvement.

\subsubsection{Dependence of Contralateral Involvement on T-Category and
Midline
Extension}\label{dependence-of-contralateral-involvement-on-t-category-and-midline-extension}

In figure~\ref{fig-model-prevalences-overall}, we plot the prevalence of
contralateral involvement of the LNLs II, III, and IV for the four
scenarios made up of the possible combinations of early and late
T-category, as well as lateralized and midline extending tumors.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_overall.pdf}}

}

\caption{\label{fig-model-prevalences-overall}Comparison of predicted
(histograms) vs observed (beta posteriors) prevalences, shown for the
contralateral LNLs II (blue), III (orange), and IV (green). The top row
shows scenarios with early T-category tumors, the bottom row for late
T-category ones. The left column depicts scenarios where the primary
tumor is clearly lateralized, the right column scenarios of tumors
extending over the mid-sagittal line. This figure illustrates the
model's ability to describe the prevalence of involvement for different
combinations of the risk factors T-category and midline extension.}

\end{figure}%

Figure~\ref{fig-model-prevalences-overall} shows that the model is
capable of accurately accounting for the most important risk factors,
i.e.~T-category and midline extension. As observed in the data, the
model predicts that the prevalence of contralateral LNL II involvement
increases from 7.1\% for early T-category lateralized tumors to 39.17\%
when the tumor is of advanced T-category and crosses the mid-sagittal
line. Similarly, the prevalence of contralateral LNL III involvement
increases from around 1.55\% for early T-category lateralized tumors to
almost 14.25\% when the tumor is of advanced T-category and crosses the
mid-sagittal line.

\subsubsection{Influence of Upstream Involvement on Contralateral
Metastasis}\label{influence-of-upstream-involvement-on-contralateral-metastasis}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_upstream.pdf}}

}

\caption{\label{fig-model-prevalences-upstream}The influence of the
upstream LNL II's involvement on the prevalence of contralateral level
III for the four combinations of tumor lateralization (lateralized or
extending over midline) and T-ctageory (early or advanced). Our model
predictions (histograms) are plotted against the observations in the
data (beta posteriors).}

\end{figure}%

Figure~\ref{fig-model-prevalences-upstream} illustrates how rarely the
contralateral LNL III harbors metastases when its upstream level II is
healthy. Also, the model can capture this correlation rather well, with
the exception of early T-category tumors that extend over the
mid-sagittal line. As in figure~\ref{fig-model-prevalences-overall} and
figure~\ref{fig-model-prevalences-with-ipsi}, these cases are relatively
rare resulting in a broad distribution over the true prevalence.

\subsubsection{Correlation between Ipsi- and Contralateral
Involvement}\label{correlation-between-ipsi--and-contralateral-involvement}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_with_ipsi.pdf}}

}

\caption{\label{fig-model-prevalences-with-ipsi}Comparison of the
computed and observed prevalences for scenarios that illustrate the
model's capability of accounting for the correlation between ipsi- and
contralateral involvement. We show three scenarios where we consider the
joint involvement of contralateral LNL II together with different
ipsilateral involvements: 1) the ipsilateral neck shows no involvement
in blue (LNLs I to V are healthy, LNL VII is unspecified because data on
it is missing for some patients), 2) where ipsilateral LNL II is
involved in orange (LNLs I, III, IV, and V are healthy), and 3) where
ipsilateral LNLs II and III are involved in orange (LNLs I, IV, and V
are healthy). These two scenarios are plotted for all combinations of
T-category (early in top row, advanced in bottom row) and tumor
lateralization (lateralized in left column, extending over mid-sagittal
line in the right column).}

\end{figure}%

In figure~\ref{fig-model-prevalences-with-ipsi} we display the model's
ability to capture the correlation between ipsi- and contralateral
involvement. Shown are marginals of the joint distribution where we
consider contralateral LNL II involvement with simulaneous involvement
of different ipsilateral LNLs. The model has no direct connections
between the two sides and thus no adjustable parameter to specifically
quantify correlations between ipsi- and contralateral involvement.
Metastases in the two sides of the neck are only correlated via the time
of diagnosis. Nevertheless, the model adequately predicts that
contralateral level II involvement is rare in combination with a healthy
ipsilateral neck (green histograms). If ipsilateral LNL II is involved,
contralateral involvement is more frequent.

\section{Results: Prediction of Risk for Occult
Disease}\label{sec-results-risk}

For a clinical application of the model we are interested in the risk of
occult metastases rather than the observed patterns of lymph node
involvement. We want to estimate the risk for occult disease in
clinically negative LNLs, given the patient's individual diagnosis. In
terms of our model, the diagnosis consists of the T-category, the
lateralization of the tumor (if it extend over the mid-sagittal plane),
and which LNLs are clinically involved based on imaging and possibly
fine needle aspiration (FNA). For the results shown in this section, we
assume that involvement of a LNL is clinically diagnosed through imaging
with a sensitivity of 81\% and a specificity of 76\%
\citep{de_bondt_detection_2007}. For FNA we assume a specificity of 98\%
and a sensitivity of 80\%. This amounts to the assumption that lymph
node involvement confirmed by FNA is almost certainly true involvement,
i.e.~there are almost no false positive diagnoses.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_risks.pdf}}

}

\caption{\label{fig-model-risks}Histograms over the predicted risk of
occult involvement in contralateral LNL II (left), III (middle), and IV
(right), shown for various combinations of T-category, tumor
lateralization, and clinical LNL diagnoses. All LNLs not explicitly
mentioned in the legend, including the LNL for which the risk of occult
disease was computed, where assumed to be clinically negative
(specificity 76\%, sensitivity 81\%).}

\end{figure}%

\subsection{Contralateral LNL II}\label{contralateral-lnl-ii}

Figure~\ref{fig-model-risks} (left panel) shows the predicted risk of
occult disease in contralateral LNL II. The most important variable
impacting the prediction for contralateral level II involvement in our
model is the tumor's lateralization. A patient with a clearly
lateralized early T-category tumor and clinically detected metastases in
ipsilateral LNL II is predicted to have a 1-2\% risk for occult disease
in contralateral LNL II (green histogram). For a cN0 patient with a
clearly lateralized early T-category tumor but \emph{with} mid-sagittal
extension, the risk increases to 7.58\% (orange histogram).

Also advanced T-category increases the risk of occult disease but plays
a lesser role. Considering the scenario of an early T-category tumor
that crosses the midline and an ipsilateral neck where the LNLs II and
III are clinically involved, the risk for occult contralateral LNL II
disease is around 9.06\% (red histogram). For the same scenario and an
advanced T-category tumor, the risk increases to 11.25\% (purple
histogram).

Lastly, the predicted risk of occult contralateral LNL II involvement
also increases with the degree of ipsilateral involvement. When an early
T-category patient with a tumor that extends over the mid-sagittal line
presents with a clinically N0 ipsilateral neck instead of involvement in
the levels II and III, the risk is 7.58\% instead of 9.06\%.

Taken together, for lateralized tumors, the risk of occult contralateral
involvement remains low even for patients with advanced T-category
tumors and ipsilateral involvement of LNLs II and III (blue histogram),
making midline extension the main risk factor. However, advanced
T-category and severe ipsilateral involvement may still considerably
increase the predicted risk for contralateral involvement.

\subsection{Contralalteral LNL III}\label{contralalteral-lnl-iii}

As shown in the center subplot of figure~\ref{fig-model-risks}, the risk
for occult disease in the contralateral LNL III may only cross a 5\%
threshold if the upstream level II is clinically involved. Even for
advanced T-category tumors extending over the midline with extensive
ipsilateral involvement (green histogram), the risk of occult metastases
in contralateral LNL III is only 2\% if the upstream LNL II is
clinically negative. For an advanced T-category tumors extending over
the midline with clinical involvement of both ispilateral and
contralateral LNL II involvement, the predicted risk in contralateral
LNL III increases to 4-5\% (blue histogram). If the clinical diagnosis
of the contralateral upstream level II is pathologically confirmed by
FNA the risk increases to 6.94\% (red histogram). This is because the
possibility that the clinical involvement in contralateral LNL II is a
false positive diagnoses is almost excluded. If contralateral LNL II is
involved, the predicted risk in contralateral LNL III is 5.58\% even for
lateralized tumors (orange histogram).

\subsection{Contralateral LNL IV}\label{contralateral-lnl-iv}

Our model's prediction for the risk in contralateral LNL IV is below 1\%
if the upstream LNL III is clinically negative, that even in the most
extreme case of an advanced T-category tumor extending over the
mid-sagittal line with metastases in the ipsilateral LNLs II-IV and the
contralateral level II (green histogram). If additionally contralateral
LNL III is clinically involved, the risk increases to 2.81\% (blue
histogram). Only when the involvement of the upstream level III is
pathologically confirmed using FNA, we predict a substantially higher
risk of 5.68\% (orange histogram). This is because the prior probability
for contralateral LNL III and IV involvement is so low that, even given
a clinically positive diagnosis in LNL III, it is likely that the
involvement in LNL III is a false possitive diagnosis both LNLs III and
IV are negative. This possibility, however, is virtually excluded when
the involvement in contralateral level III is confirmed by FNA.

\subsection{Contralateral LNLs I, V, and
VII}\label{contralateral-lnls-i-v-and-vii}

Even for advanced T-category patients whose tumor crosses the midline
and who present with extensive ipsi- and contralateral clinical
involvement, the risk for involvement in the LNLs I and VII never
exceeds 3\%. Contralateral LNL V also shows only a very low risk for
occult disease. Only in the extreme case of severe contralateral
involvement including LNL IV, confirmed by FNA, the predicted risk is
5.49\%.

\section{Discussion}\label{sec-discussion}

\subsection{Summary}\label{summary}

In this work we present a formalism to model the ipsi- and contralateral
lymph node involvement in oropharyngeal SCC patients. An ipsilateral
model has been previously developed and published
\citep{ludwig_hidden_2021, ludwig_modelling_2023}. Based on this, we
introduce an extension that leaves the ipsilateral model unchanged, but
extends it to the contralateral side in an intuitive and comprehensible
manner. The model parameters are learned from a dataset of 833 patients
from four institutions in whom ipsilateral and contralateral involvement
was reported per per patient and LNL. The model takes the clincally
diagnosed involvement of the LNLs and the primary tumor's T-category and
lateralization into account when predicting the personalized risk for
occult disease in any LNL of interest. Owing to the relatively few
parameters, the model is highly interpretable and every parameter can be
intuitively explained, while still being able to accurately describe the
data on both ipsilateral and contralateral nodal involvement observed in
the dataset.

To the best of our knowledge, this is the most comprehensive and
detailed model of lymphatic tumor progression in oropharyngeal SCC
available. The underlying data as well as the code is publicly
available. While there was work on regional tumor spread, these models
were conceptually different, more limited in the LNLs for which they
make predictions, and not trained with real patient data
\citep{benson_markov_2006, jung_development_2017}.

\subsection{Implications for Contralateral Elective Nodal
Treatment}\label{implications-for-contralateral-elective-nodal-treatment}

The predictions of this model are being used to inform a clinical trial
on volume deescalation at the University Hospital Zurich
{[}clinicaltrials.gov, NCT06563362 (DeEscO){]}. When accepting a 5\%
risk of occult disease in any given LNL, the model suggests the
following contralateral elective irradiation (assuming the respective
LNL is clinically healthy):

\begin{itemize}
\tightlist
\item
  For patients with lateralized tumors and without clinical involvement
  of contralateral LNLs, unilateral radiotherapy is performed. No
  contralateral LNLs are electively irradiated, irrespective of
  T-category and ipsilateral involvement.
\item
  For patients without clinical involvement of contralateral LNLs but
  with tumors extending over the midline, only LNL II is electively
  irradiated.
\item
  LNL III is irradiated when LNL II is involved, irrespective of
  lateralization, T-category, and ipsilateral involvement. When the
  upstream LNL II is clinically negative, LNL III is not irradiated,
  except for the unlikely situation that contralateral LNL IV is
  involved.
\item
  Irradiate LNL IV only when the upstream level III is confirmed to be
  involved.
\item
  LNL V is not electively irradiated in most patients. Only in the most
  extreme cases, i.e., for advanced T-category tumors with midline
  extension and confirmed involvement of contralateral LNLs II to IV,
  LNL V may be irradiated.
\item
  LNLs I and VII are not irradiated unless clinically involved.
\end{itemize}

In essence, this is a summary of the results in
section~\ref{sec-results-risk}. However, it should be considered in
light of the limitations outlined in
section~\ref{sec-discussion-limitations} below.

\subsection{Limitations and Future
Work}\label{sec-discussion-limitations}

\subsubsection{T-Category Dependence}\label{t-category-dependence}

The model has only one parameter, \(p_\text{adv.}\), to describe
differences in the patterns of involvement between early and advanced
T-category. Higher involvement for advanced T-category is described by
evolving the state of involvement overr a larger number of time-steps,
while the probability of spread per time step, here the \(b_v\)
parameters, remains constant. Overall, the model captures the
differences between early and advanced T-category very well as seen in
Figure~\ref{fig-model-prevalences-overall}. However, for some situations
the observed difference between early vs advanced T-category in the data
is greater or smaller than what our model predicts. This has previously
been discussed for the involvement of ipsilateral LNL I
\citep{ludwig_modelling_2023}. In the context of contralateral spread,
the prevalence of midline extension is overestimated for early
T-category tumors and underestimated for advanced T-category tumors,
which is further discussed in section~\ref{sec-prevalence-midext}.

\subsubsection{Sensitivity and
Specificity}\label{sensitivity-and-specificity}

As discussed in section~\ref{sec-methods-consensus} (and in more detail
in section~\ref{sec-consensus}), we assumed for model training that the
consensus across all diagnostic modalities represents the true state
\(X_v\) of lymph node involvement. This is appropriate for
pathologically confirmed involvement but is an approximation for
clinically diagnosed involvement, as the latter is - by definition -
incapable of detecting occult disease. In principle, the model can be
trained while differentiating between pathologically confirmed and
clinically diagnosed involvement, using a lower sensitivity and
specificity for clinical diagnosis. This was not done here for two
reasons: 1) this simplification allowed us to compare the observed
prevalence of involvement to the model's prediction, and thereby
investigate the model's capability to describe data with few
interpretable parameters; and 2) the literature values for sensitivity
and specificity of the diagnostic modalities are inconsistent with some
aspects of the observed data. Future work may aim at developing novel
concepts to rigorously differentiating between pathologically confirmed
and clinically diagnosed involvement.

\subsubsection{Tumor Subsites}\label{tumor-subsites}

Within the cohort of patients with oropharyngeal tumors, subgroups can
be identified whose tumors are located in different subsites, e.g.~at
the base of the tongue or on the tonsils. The lymphatic metastatic
spread patterns of these subgroups may differ slightly. If so, the model
could benefit from more fine-grained information about the precise tumor
subsite for its training and prediction. We have conducted preliminary
work to show that a mixture model may be a promising approach to capture
the subsite specific spread patterns \citep[cite ICCR
paper,][]{zora231470}. The additionally represents an approach to extend
the model towards tumor locations in the oral cavity, hypopharynx, and
larynx.

\section{Acknowledgement}\label{acknowledgement}

This work was supported by:

\begin{itemize}
\tightlist
\item
  the Clinical Research Priority Program ``Artificial Intelligence in
  Oncological Imaging'' of the University of Zurich
\item
  the Swiss Cancer Research Foundation under grant number KFS
  5645-08-2022
\end{itemize}

\section{Consensus on most likely involvement}\label{sec-consensus}

The consensus on the most likely involvement state of a LNL was formed
as follows: Suppose the involvement status \(X_v\) of LNL \(v\) was
assessed using different diagnostic modalities
\(\mathcal{O} = \{ \text{MRI}, \text{CT}, \text{pathology}, \ldots \}\),
each characterized by their own pair of sensitivity and specificity
values \(s_N^{\mathcal{o}}\) and \(s_P^{\mathcal{o}}\), with
\(\mathcal{o} \in \mathcal{O}\). These values are tabulated in
table~\ref{tbl-spec-sens}. Then we have \(|\mathcal{O}|\) observations
\(z_v^{\mathcal{o}} \in \left[ 0, 1 \right]\), where 0 stands for
``healthy'' and 1 for ``involved''. We can then compute the most likely
true involvement \(X_v\) using the likelihood function

\[
\begin{aligned}
\ell \left( X_v \mid \{ z_v^{\mathcal{o}} \}_{\mathcal{o} \in \mathcal{O}} \right) = \prod_{\mathcal{o} \in \mathcal{O}}
\left( 1 - X_v \right) \cdot &\left[ z_v^{\mathcal{o}} \cdot \left( 1 - s_P^{\mathcal{o}} \right) + \left( 1 - z_v^{\mathcal{o}} \right) \cdot s_P^{\mathcal{o}} \right] \\
+ X_v \cdot &\left[ z_v^{\mathcal{o}} \cdot s_N^{\mathcal{o}} + \left( 1 - z_v^{\mathcal{o}} \right) \cdot (1 - s_N^{\mathcal{o}}) \right]
\end{aligned}
\]

We now assume the true state \(X_v\) to take on the value 1 if
\(\ell \left( X_v = 1 \mid \ldots \right) > \ell \left( X_v = 0 \mid \ldots \right)\)
and 0 otherwise. For example, if we have \(z_\text{II}^\text{CT} = 0\)
and \(z_\text{II}^\text{MRI} = 1\) we would compute the following
likelihoods:

\[
\begin{aligned}
\ell \left( X_\text{II} = 1 \mid z_\text{II}^\text{CT} = 0, z_\text{II}^\text{MRI} = 1 \right) &= \left( 1 - s_N^\text{CT} \right) \cdot s_N^\text{MRI} = 15.39\% \\
\ell \left( X_\text{II} = 0 \mid z_\text{II}^\text{CT} = 0, z_\text{II}^\text{MRI} = 1 \right) &= s_P^\text{CT} \cdot \left(1 - s_N^\text{MRI}\right) = 14.44\%
\end{aligned}
\]

In this example, we would thus assume the true state to be involved
(\(X_\text{II} = 1\)).

This method of computing a consensus also ensures that the pathology
reports always override any conflicting clinical diagnosis, due to
pathology's high sensitivity and specificity.

\begin{longtable}[]{@{}lrr@{}}
\caption{Specificity and sensitivity values from the literature
\citep{de_bondt_detection_2007, kyzas_18f-fluorodeoxyglucose_2008}.}\label{tbl-spec-sens}\tabularnewline
\toprule\noalign{}
Modality & Specificity & Sensitivity \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Modality & Specificity & Sensitivity \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
CT & 76\% & 81\% \\
PET & 86\% & 79\% \\
MRI & 63\% & 81\% \\
FNA & 98\% & 80\% \\
pathology & 100\% & 100\% \\
\end{longtable}

\section{Contralateral Prevalence of
Involvement}\label{contralateral-prevalence-of-involvement}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\begin{longtable}[]{@{}llllllllllll@{}}

\caption{\label{tbl-data-strat}Contralateral involvement depending on
whether the primary tumor extends over the mid-sagittal line, the
T-category, and whether the ipsilateral LNL III was involved or
healthy.}

\tabularnewline

\caption{}\label{T_fde0d}\tabularnewline
\toprule\noalign{}
T-cat. & ipsi & Mid. ext. & \multicolumn{2}{l}{%
I} & \multicolumn{2}{l}{%
II} & \multicolumn{2}{l}{%
III} & \multicolumn{2}{l}{%
IV} & total \\
& & & n & \% & n & \% & n & \% & n & \% & n \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
T-cat. & ipsi & Mid. ext. & \multicolumn{2}{l}{%
I} & \multicolumn{2}{l}{%
II} & \multicolumn{2}{l}{%
III} & \multicolumn{2}{l}{%
IV} & total \\
& & & n & \% & n & \% & n & \% & n & \% & n \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
early & 0 & False & 0 & 1 & 0 & 0 & 0.00 & 1.16 & 0.00 & 0.00 & 86 \\
early & 0 & True & 0 & 1 & 1 & 0 & 0.00 & 10.00 & 10.00 & 0.00 & 10 \\
early & 0 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 12 \\
early & 1 & False & 1 & 11 & 2 & 1 & 0.53 & 5.82 & 1.06 & 0.53 & 189 \\
early & 1 & True & 1 & 2 & 0 & 0 & 11.11 & 22.22 & 0.00 & 0.00 & 9 \\
early & 1 & nan & 0 & 3 & 0 & 1 & 0.00 & 9.68 & 0.00 & 3.23 & 31 \\
early & ≥ 2 & False & 1 & 15 & 3 & 3 & 0.96 & 14.42 & 2.88 & 2.88 &
104 \\
early & ≥ 2 & True & 0 & 3 & 4 & 1 & 0.00 & 30.00 & 40.00 & 10.00 &
10 \\
early & ≥ 2 & nan & 0 & 3 & 0 & 0 & 0.00 & 13.64 & 0.00 & 0.00 & 22 \\
advanced & 0 & False & 0 & 2 & 0 & 0 & 0.00 & 5.88 & 0.00 & 0.00 & 34 \\
advanced & 0 & True & 0 & 3 & 0 & 0 & 0.00 & 12.50 & 0.00 & 0.00 & 24 \\
advanced & 0 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 3 \\
advanced & 1 & False & 0 & 3 & 0 & 0 & 0.00 & 4.55 & 0.00 & 0.00 & 66 \\
advanced & 1 & True & 1 & 18 & 5 & 1 & 1.64 & 29.51 & 8.20 & 1.64 &
61 \\
advanced & 1 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 9 \\
advanced & ≥ 2 & False & 4 & 16 & 7 & 4 & 5.26 & 21.05 & 9.21 & 5.26 &
76 \\
advanced & ≥ 2 & True & 3 & 46 & 18 & 8 & 3.80 & 58.23 & 22.78 & 10.13 &
79 \\
advanced & ≥ 2 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 8 \\

\end{longtable}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

\section{Prevalence of Midline Extension}\label{sec-prevalence-midext}

\begin{figure}

\centering{

\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_midext.pdf}

}

\caption{\label{fig-model-prevalences-midext}Comparing the predicted
(histograms) and observed (lines depicting beta posteriors) prevalence
of midline extension for early (blue) and late (orange) T-category.
While the prevalence is predicted correctly when marginalizing over
T-category, the model cannot capture the degree of separation observed
in the data. Since the tumor's midline extension is virtually always
part of the diagnosis and hence \emph{given} when predicting a patient's
risk, we do not consider this discrepancy a major issue.}

\end{figure}%

In figure~\ref{fig-model-prevalences-midext}, we plot the prevalence of
midline extension in the data versus our model's prediction. It is
obvious the model cannot match the large spread between early and
advanced T-category seen in the data. This is because to achieve that,
it would need to increase the advanced T-category patient's prior
distribution over diagnosis times and at the same time reduce the
probability of the tumor to cross the midline during a time-step. But
since the time-priors parameter is also coupled with the spread
probabilities among the LNLs, the model does not have that freedom.

However, we do not consider this discrepancy a major limitation of the
model: We will not realistically be interested in the probability of
midline extension, as it is always possible to assess it with high
certainty. That is also the reason why we initially modelled the midline
extension \emph{not} as a random variable, but as a global risk factor
that would have been turned on or off from the onset of a patient's
disease evolution. This, however, lead to overly high risks for
contralateral involvement in advanced T-category patients with midline
extension, because then the model assumes an increased spread to the
contralateral side from the onset of the disease. Which is probably not
true in a majority of those cases. Thus, treating it as a random
variable that only becomes true during a patient's disease evolution
resulted in a better description of the data.

Formally, the wrong prediction of midline extension prevalence makes
little difference, since it is always given: Instead of
\(P\left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right)\),
we typically compute
\(P\left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c}, \epsilon \right)\),
which does not suffer from the wrong probability of midline extension,
as the distribution over hidden states is renormalized:

\[
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c}, \epsilon \right) = \frac{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right) P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right)}{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c}, \epsilon \right)}
\]

Note that a distribution over \(\epsilon\) appears both in the
enumerator and the denominator, which largely cancel each other, leaving
only the midline extension's effect on the distribution over hidden
states in the prediction.

Also, the discrepancy in midline extension prevalence between early and
advanced T-category is particularly pronounced in oropharyngeal SCC
patients. For example, in oral cavity SCC, the midline extension only
increases from 15.4\% (20 out of 130) to 33.3\% (13 out of 39).

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript-preview.html}{Article
Notebook}}

  \bibliography{references.bib}

\end{document}