### 2024-11-26 manuscript.tex
rerun pipeline

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-mathphys-num,
]{sn-jnl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\usepackage{multirow}
\usepackage{centernot}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Probabilistic Model of Bilateral Lymphatic Spread in Head and
Neck Cancer]{Probabilistic Model of Bilateral Lymphatic Spread in Head
and Neck Cancer}

% author setup
\author*[1,2]{\fnm{Roman} \sur{Ludwig}}\email{roman.ludwig@usz.ch}\author[1,2]{\fnm{Yoel Perez} \sur{Haas}}\email{yoel.perezhaas@usz.ch}\author[1,2]{\fnm{Jan} \sur{Unkelbach}}\email{jan.unkelbach@usz.ch}
% affil setup
\affil[1]{\orgdiv{Department of Physics}, \orgname{University of
Zurich}}
\affil[2]{\orgdiv{Radiation Oncology}, \orgname{University Hospital
Zurich}}

% abstract

\abstract{Purpose: According to current guidelines for elective nodal
irradiation of oropharyngeal squamous cell carcinoma (OPSCC) patients,
large parts of the contralateral lymphatic system are included in the
elective clinical target volume (CTV-N), even for lateralized tumors
without clinical lymph node involvement in the contralateral neck. In
this work, we present a probabilistic model for bilateral lymphatic
tumor progression in OPSCC to predict the personal risk for occult
disease in any lymph node level (LNL), given the patient's clinical
lymph node involvement, T-stage, and lateralization of the primary
tumor.

Methods: We extend a hidden markov model for lymphatic tumor
progression, which was previously developed for ipsilateral lymph node
involvement, to the contralateral neck. The model represents each of the
LNLs I, II, III, IV, V, and VII of both sides of the neck as a hidden
binary random variable with states healthy and involved. LNLs are
connected to the tumor and among each other via arcs that correspond to
spread probabilities. These spread probability rates are learned via
Markov chain Monte Carlo (MCMC) sampling from a dataset of 833 OPSCC
patients.

Results: The model is able to describe the data on lymph node
involvement well with a small number of interpretable parameters.
Midline extension of the primary tumor is the main risk factor for
contralateral involvement. In addition, the risk of occult metastases in
contralateral lymph node level increases with more advanced T-stage and
more severe ipsilateral involvement. The probability of involvement in
contralateral level III is very low if the upstream level II is
clinically negative. Similarly, the probability of involvement in
contralateral level IV is very low if the upstream level III is
clinically negative.

Conclusions: The model may guide personalized volume reduction of the
elective CTV-N. It suggests that the contralateral neck may be spared
from elective irradiation for lateralized tumors not crossing the
midline. In patients with primary tumors crossing the midline but
clinically negative contralateral neck, the contralateral elective
irradiation may be limited to level II.}

% keywords

\begin{document}
\maketitle

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\section{Introduction}\label{introduction}

In treating head and neck squamous cell carcinomas (HNSCC) with
radiotherapy or surgery, both the primary tumor and clinically detected
lymph node metastases are targeted. Current guidelines also include
large portions of the neck in the elective clinical target volume
(CTV-N)
\citep{gregoire_ctbased_2003, gregoire_delineation_2014, gregoire_delineation_2018, eisbruch_intensitymodulated_2002, biau_selection_2019, chao_determination_2002, vorwerk_guidelines_2011, ferlito_elective_2009}
to mitigate the risk of regional recurrences from untreated microscopic
disease undetectable by in-vivo imaging modalities such as computed
tomography (CT), magnetic resonance imaging (MRI), or positron emission
tomography (PET). However, this approach must balance minimizing the
risk of occult disease in the lymphatic drainage region against the
toxicity of unnecessarily irradiating healthy tissue.

These CTV-N guidelines rely on anatomically defined lymph node levels
(LNLs) \citep{gregoire_delineation_2014} and the overall prevalence of
lymph node metastases within these levels. They often recommend
extensive irradiation of both sides of the neck. However, the general
prevalence of metastasis in a given LNL does not correspond to an
individual patient's risk of occult disease in that region, which
depends on their specific state of tumor progression. For example, a
patient with no clinically detectable nodal disease (cN0) who has a
small, clearly lateralized T1 tumor would receive the same contralateral
CTV-N as a patient with significant ipsilateral nodal involvement and an
advanced tumor crossing the mid-sagittal plane: Both patients receive
elective irradiation of the contralateral LNLs II, III, and IVa
\citep{biau_selection_2019}.

To better quantify individualized risk of occult disease, we previously
developed an intuitive probabilistic hidden Markov model (HMM)
\citep{ludwig_hidden_2021, ludwig_modelling_2023}, originally based on a
conceptually similar a Bayesian network model
\citep{pouymayou_bayesian_2019}. However, these models have been limited
to predicting ipsilateral nodal involvement. This work extends the model
to include contralateral risk predictions, enabling more personalized
radiation volume recommendations. By identifying patients with low
contralateral risk, the model could guide reductions in contralateral
CTV-N, thereby decreasing radiation-induced toxicity and improving
quality of life.

The main contributions of this paper are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Section section~\ref{sec-data} presents a multi-centric dataset on
  lymph node involvement in 833 OPSCC patients, identifying key risk
  factors for contralateral lymph node involvement and outlining
  requirements for a bilateral model extension
  (section~\ref{sec-requirements}).
\item
  Section section~\ref{sec-ext-to-contra} introduces a bilateral HMM
  that incorporates primary tumor lateralization, T-category, and
  clinical involvement as risk factors for contralateral involvement.
  Model training and computational experiments are described in
  section~\ref{sec-methods}.
\item
  Section section~\ref{sec-results} demonstrates the model's ability to
  replicate observed contralateral lymph node involvement patterns and
  estimates occult disease risk for typical patients. Implications for
  volume-deescalated radiotherapy are discussed in
  section~\ref{sec-discussion}.
\end{enumerate}

\section{Data on Lymphatic Progression Patterns}\label{sec-data}

To develop models for lymphatic tumor progression for all relevant LNLs,
including contralateral regions, we compiled a detailed dataset of 833
patients with newly diagnosed oropharyngeal squamous cell carcinomas
(OPSCC) \citep{ludwig_dataset_2022, ludwig_multicentric_2024}. The
dataset includes lymph node involvement per LNL for each patient in
tabular form, along with primary tumor and patient characteristics such
as T-category, subsite, primary tumor lateralization, and HPV p16
status. Patient records were collected from four institutions, and an
overview of patient characteristics is provided in
table~\ref{tbl-data-overview}.

Data from Inselspital Bern (ISB) and Centre Léon Bérard (CLB) consist
exclusively of patients who underwent neck dissections. In contrast, the
majority of patients from the University Hospital Zürich (USZ) and the
Hospital Vall d'Hebron (HVH) were treated with definitive radiotherapy.
Since surgical treatment is more common for early T-category patients,
ISB and CLB datasets include a higher proportion of these cases compared
to USZ and HVH. For 83 patients in the CLB dataset, the primary tumor's
lateralization was not reported.

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\begin{longtable}[]{@{}lllllll@{}}

\caption{\label{tbl-data-overview}Overview over the five datasets from
four different institutions used to train and evaluate our model. Here,
we briefly characterize the total number of OPSCC patients from the
respective institution, their median age, what proportion received some
form of neck dissection, the N0 portion of patients, what percentage
presented with early T-category, and the prevalence of primary tumor
midline extension. For a much more detailed look at the data, visit
\href{https://lyprox.org}{lyprox.org}.}

\tabularnewline

\caption{}\label{T_561f9}\tabularnewline
\toprule\noalign{}
Institution & Total & Age (median) & Neck Dissection & N0 & Early T-Cat.
& Mid. Ext. \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Institution & Total & Age (median) & Neck Dissection & N0 & Early T-Cat.
& Mid. Ext. \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Centre Léon Bérard & 325 & 60 & 100\% & 19\% & 69\% & 18\% \\
Inselspital Bern & 74 & 61 & 100\% & 18\% & 66\% & 14\% \\
University Hospital Zurich & 287 & 66 & 26\% & 18\% & 52\% & 31\% \\
Vall d\textquotesingle Hebron Barcelona Hospital & 147 & 58 & 5\% & 21\%
& 34\% & 34\% \\

\end{longtable}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\subsection{Consensus on Involvement Status}\label{sec-data-consensus}

Pathological involvement is available only for surgically treated
patients and for the levels that were dissected. For non-surgical
patients, involvement status is determined clinically, i.e.~using
imaging. To ensure uniformity in analysis, diagnostic information was
synthesized into a consensus decision for each patient and LNL. This
consensus reflects the most likely state of involvement and accounts for
the sensitivity and specificity of various diagnostic modalities, as
reported in the literature
\citep{debondt_detection_2007, kyzas_18ffluorodeoxyglucose_2008}.

The consensus process is detailed in section~\ref{sec-consensus}.
Briefly, pathological findings from neck dissections are treated as the
gold standard, overriding any conflicting clinical diagnoses. For levels
not dissected, PET-CT is typically the primary source for determining
the most likely state of involvement.

\subsection{Data Availability}\label{data-availability}

The complete dataset, including additional patients with tumors in
primary sites other than the oropharynx, is publicly accessible. It can
be \href{https://lyprox.org/patients/dataset}{downloaded from LyProX},
where it is also available for interactive exploration,
\href{https://github.com/rmnldwg/lydata}{from GitHub}, or
\href{https://zenodo.org/search?q=lydata}{from Zenodo}. Additional
details about the datasets and data format are provided in designated
Data-in-Brief publications
\citep{ludwig_dataset_2022, ludwig_multicentric_2024}. These
publications do not include the latest data from HVH, which will be
described in a future publication.

\subsection{Patterns of Contralateral Involvement}\label{sec-data-strat}

The datasets enable analysis of correlations between LNL involvement and
key risk factors. In figure~\ref{fig-data-strat}, we illustrate the
prevalence of contralateral LNL involvement, stratified by T-category,
the number of ipsilaterally involved LNLs, and whether the tumor extends
across the mid-sagittal plane.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_data_strat.pdf}}

}

\caption{\label{fig-data-strat}Contralateral involvement stratified by
T-category (top left panel), the number of metastatic LNLs ipsilaterally
(top right panel), and whether the primary tumor extended over the
mid-sagittal line or was clearly lateralized (bottom left panel).
Additionally, in the bottom right, we show that even for lateralized
tumors, T-category and ipsilateral involvement are correlated with
contralateral metastasis by comparing the contralateral involvement
prevalence for selected scenarios that vary in their T-category and
ipsilateral involvement extent.}

\end{figure}%

\subsubsection{Midline Extension}\label{midline-extension}

The bottom left panel of figure~\ref{fig-data-strat} shows that tumors
crossing the mid-sagittal plane have a substantially higher prevalence
of contralateral involvement compared to clearly lateralized tumors.
This aligns with the anatomy of the head and neck lymphatic system,
which is symmetric, with no major lymph vessels crossing the midline.
Interstitial fluids from the primary tumor, presumed to carry malignant
cells, can merely diffuse to contralateral lymphatic vessels over short
distances. Thus, contralateral spread is more likely when the tumor
approaches or crosses the midline.

\subsubsection{T-Category}\label{t-category}

The top left panel shows a correlation between T-category and
contralateral involvement, reflecting T-category's role as a surrogate
for the time elapsed between disease onset and diagnosis. Advanced
T-category tumors (e.g., T4) generally represent a longer disease
progression timeline, providing more opportunity for metastatic spread
compared to smaller tumors (e.g., T1).

\subsubsection{Ipsilateral Involvement}\label{ipsilateral-involvement}

The top right panel reveals a positive correlation between ipsilateral
and contralateral metastases. Extensive ipsilateral involvement likely
indicates a longer or faster disease progression. Additionally, it has
been hypothesized that bulky ipsilateral nodal disease may reroute
lymphatic drainage toward the contralateral side, potentially
facilitating contralateral metastasis.

\subsubsection{Interplay of Risk
Factors}\label{interplay-of-risk-factors}

Midline extension, T-category, and ipsilateral involvement are
interrelated risk factors for contralateral metastasis. For instance,
45.56\% of advanced T-category tumors exhibit midline extension compared
to 6.13\% of early T-category tumors. While the higher fraction of
midline extensions in advanced T-category patients partially accounts
for the higher contralateral metastasis rates, T-category itself and
ipsilateral involvement also play an additional role.

The bottom right panel of figure~\ref{fig-data-strat} isolates patients
with lateralized tumors that do not cross the midline, further
demonstrating these relationships. Among early T-category patients with
no ipsilateral nodal involvement (levels I-V), only 1.85\% (2 of 108
patients) show involvement in contralateral level II. This proportion
increases to 6.73\% (14 of 208) if ipsilateral level II is involved, to
14.29\% (10 of 70) if ipsilateral levels II and III are involved, and
further to 39.68\% (25 of 63) for advanced T-category tumors with
ipsilateral levels II and III involved.

\subsection{Requirements for a Bilateral Model}\label{sec-requirements}

Based on the observations in section~\ref{sec-data-strat} above, any
model predicting the risk of contralateral nodal involvement should
account for the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Midline Extension:} Tumors extending across the mid-sagittal
  plane should result in a significantly higher probability of
  contralateral metastases.
\item
  \textbf{T-Category and Disease Progression:} Advanced T-category
  should correspond to an increased risk of nodal disease. This can be
  modeled using the expected time of diagnosis, as demonstrated in a
  prior hidden Markov model \citep{ludwig_hidden_2021}.
\item
  \textbf{Ipsilateral Involvement as a Progression Marker:} The extent
  of ipsilateral involvement should provide additional information on
  the time elapsed between disease onset and diagnosis, supplementing
  the insights gained from T-category alone.
\end{enumerate}

\section{Unilateral Model for Lymphatic
Progression}\label{sec-unilateral}

This paper builds on the previously developed unilateral model for
ipsilateral lymph node involvement presented in
\citep{ludwig_modelling_2024}. Below, we briefly recap the unilateral
model to introduce the notation required for extending the framework to
a bilateral model, described in section~\ref{sec-ext-to-contra}. For
further details on the ipsilateral model, refer to earlier publications
\citep{ludwig_hidden_2021, ludwig_modelling_2024}.

We represent a patient's state of involvement at an abstract time-step
\(t\) as a vector of hidden binary random variables, where each
component corresponds to a lymph node level (LNL):

\begin{equation}\phantomsection\label{eq-state-def}{
\mathbf{X}[t] = \begin{pmatrix} X_v[t] \end{pmatrix} \qquad v \in \left\{ 1, 2, \ldots, V \right\}
}\end{equation}

Here, \(V\) is the number of LNLs the model considers. The values a
LNL's may take on are \(X_v[t] = 0\) (\texttt{False}), meaning the LNL
\(v\) is healthy or free of metastatic disease, or \(X_v[t] = 1\)
(\texttt{True}), corresponding to some form of tumor presence (i.e.,
occult or macroscopic disease). In total, there are \(2^V\) distinct
possible lymphatic involvement patterns, which we enumerate from
\(\boldsymbol{\xi}_0 = \begin{pmatrix} 0 & 0 & \cdots & 0 \end{pmatrix}\)
to
\(\boldsymbol{\xi}_{2^V} = \begin{pmatrix} 1 & 1 & \cdots & 1 \end{pmatrix}\).
Each LNL's state is observed via another binary random variable \(Z_v\)
that describes the clinical involvement of a LNL based on imaging:
\(Z_v = 0\) (\texttt{False}) indicates that the LNL \(v\) is healthy
based on clinical diagnosis, and \(Z_v = 1\) (\texttt{True}) indicates
that LNL \(v\) was classified as involved. \(X_v\) and \(Z_v\) are
connected through the sensitivity and specificity of the diagnositc
modality.

Based on this, our HMM is fully described by defining the following
three quantities:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A starting state \(\mathbf{X}[t=0]\) at time \(t=0\) just before the
  patient's tumor formed. In our case, this is always the state
  \(\boldsymbol{\xi}_0\) where all LNLs are still healthy.
\item
  The \emph{transition matrix}
  \begin{equation}\phantomsection\label{eq-trans-matrix}{
  \mathbf{A} = \left( A_{ij} \right) = \big( P \left( \mathbf{X}[t+1] = \boldsymbol{\xi}_j \mid \mathbf{X}[t] = \boldsymbol{\xi}_i \right) \big)
  }\end{equation} where the value at row \(i\) and column \(j\)
  represents the probability to transition from state
  \(\boldsymbol{\xi}_i\) to \(\boldsymbol{\xi}_j\) during the time-step
  from \(t\) to \(t+1\). Note that we prohibit self-healing, meaning
  that during a transition, no LNL may change their state from
  \(X_v[t]=1\) to \(X_v[t+1]=0\). Consequently, many elements of the
  transition matrix are zero.
\item
  Lastly, the \emph{observation matrix}
  \begin{equation}\phantomsection\label{eq-obs-matrix}{
  \mathbf{B} = \left( B_{ij} \right) = \big( P \left( \mathbf{Z} = \boldsymbol{\zeta}_j \mid \mathbf{X}[t_D] = \boldsymbol{\xi}_i \right) \big)
  }\end{equation} where in row \(i\) and at column \(j\) we find the
  probability to \emph{observe} a lymphatic involvement pattern
  \(\mathbf{Z} = \boldsymbol{\zeta}_j\), given that the true (but
  hidden) state of involvement at the time of diagnosis \(t_D\) is
  \(\mathbf{X}[t_D] = \boldsymbol{\xi}_i\).
\end{enumerate}

The transition matrix \(\mathbf{A}\) is parameterized using a directed
acyclic graph (DAG) that represents the underlying lymphatic network.
Edges from the primary tumor to an LNL are associated with a probability
\(b_v\) for direct spread to LNL \(v\) during one time step. Arcs from a
LNL \(v\) to a LNL \(r\) are parameterized with the probability rate
\(t_{vr}\) for inter-nodal progression. In this paper, we build on the
DAG shown in figure~\ref{fig-full-graph} which was obtained by
maximizing the model evidence as described in
\citep{ludwig_modelling_2024}.

\begin{figure}

\centering{

\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/static/full-graph.pdf}

}

\caption{\label{fig-full-graph}Directed acyclic graph (DAG) representing
the abstract lymphatic network in the head and neck region. Blue nodes
are the LNLs' hidden random variables, the red node represents the
tumor, and the orange square nodes depict the binary observed variables.
Red and blue arcs symbolize the probability of lymphatic spread along
that edge during one time-step. The orange arcs represent the
sensitivity and specificity of the observational modality (e.g.~CT, MRI,
pathology, \ldots).}

\end{figure}%

We can evolve the distribution of all possible hidden states from
\(\mathbf{X}[t=0] = \boldsymbol{\xi}_0\) step by step, by successively
multiplying this vector with the transition matrix \(\mathbf{A}\):
\(\mathbf{X}[t+1] = \mathbf{X}[t] \cdot \mathbf{A}\). For later use, we
define at this point a matrix \(\boldsymbol{\Lambda}\) that collects
these distributions for all considered time-steps:

\begin{equation}\phantomsection\label{eq-lambda-matrix}{
\boldsymbol{\Lambda} = P \left( \mathbf{X} \mid \mathbf{t} \right) = \begin{pmatrix}
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^0 \\
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^1 \\
\vdots \\
\boldsymbol{\pi}^\intercal \cdot \mathbf{A}^{t_\text{max}} \\
\end{pmatrix}
}\end{equation}

Where the \(k\)-th row in this matrix corresponds to the probability
distribution over hidden states after \(t=k-1\) time-steps.

At the time of diagnosis, \(0 \leq t_D \leq t_\text{max}\), we multiply
the evolved state distribution with the observation matrix
\(\mathbf{B}\) to obtain the distribution over all possible diagnoses.
However, the exact time of diagnosis, \(t_D\), is unknown; that is, we
do not know the number of time-steps over which the HMM should be
evolved. To address this, we marginalize over all possible diagnosis
times, allowing the diagnosis to occur at any time-step, albeit with
different weights. These weights are defined by a prior distribution
over \(t_D\), which can vary depending on the patient's T-category. For
example, the time-prior for early T-category patients,
\(P(t_D \mid \text{early})\), may put more weight on earlier time-steps,
reflecting -- on average -- earlier detection, compared to the prior for
advanced T-category patients, \(P(t_D \mid \text{advanced})\).

\[
P\left( \mathbf{X} \mid \text{T}x = \text{early} \right) = \sum_{t=0}^{t_\text{max}} P \left( \mathbf{X} \mid t \right) \cdot P(t \mid \text{early})
\]

In this work, we use binomial distributions
\(\mathfrak{B} \left( t_D, p_{\text{T}x} \right)\) as time-priors which
have one free parameter \(p_{\text{T}x}\) for each group of patients we
differentiate based on T-category. Also, we fix \(t_\text{max} = 10\),
which means that the expected number of time-steps from the onset of a
patient's disease to their diagnosis is
\(\mathbb{E}\left[ t_D \right] = 10 \cdot p_{\text{T}x}\).

\subsection{Likelihood Function of the Unilateral
Model}\label{likelihood-function-of-the-unilateral-model}

The likelihood function for a patient to present with a diagnosis
\(\boldsymbol{\zeta}_i\) a T-category \(\text{T}x\) tumor can now be
written as:

\begin{equation}\phantomsection\label{eq-single-patient-llh}{
\ell = P \left( \mathbf{Z} = \boldsymbol{\zeta}_i \mid \text{T}x \right) = \sum_{t=0}^{t_\text{max}} \left[ \boldsymbol{\xi}_0 \cdot \mathbf{A}^t \cdot \mathbf{B} \right]_i \cdot P \left( t \mid \text{T}x \right)
}\end{equation}

With \(\left[ \ldots \right]_i\) we denote the \(i\)-th component of the
vector in the square brackets. Note that it is also possible to account
for missing involvement information: If a diagnosis (like fine needle
aspiration (FNA)) is only available for a subset of all LNLs, we can sum
over all those possible complete observed states
\(\boldsymbol{\zeta}_j\) that match the provided diagnosis.

The single-patient likelihood \(\ell\) in
equation~\ref{eq-single-patient-llh} depends on the spread parameters
shown in figure~\ref{fig-full-graph} via the transition matrix
\(\mathbf{A}\) and on the binomial parameters \(p_{\text{T}x}\) via
time-priors. In this work, we will only differentiate between ``early''
(T1 \& T2) and ``advanced'' (T3 \& T4) T-categories. Therefore, the
parameter space of the unilateral model is:

\begin{equation}\phantomsection\label{eq-param-space}{
\boldsymbol{\theta} = \left( \left\{ b_v \right\}, \left\{ t_{vr} \right\}, p_\text{early}, p_\text{adv.} \right) \quad \text{with} \quad \genfrac{}{}{0pt}{2}{v\leq V}{r\in\operatorname{pa}(v)}
}\end{equation}

And it is our goal to infer optimal parameter values of from a given
dataset \(\mathcal{D}\) (consisting of diagnoses and T-categories) of
OPSCC patients. The likelihood to observe this cohort of \(N\) patients,
given a set of parameters \(\boldsymbol{\theta}\) is simply the product
of their individual likelihoods as defined in
equation~\ref{eq-single-patient-llh}. For numerical reasons, we
typically compute the data likelihood in log space:

\begin{equation}\phantomsection\label{eq-log-likelihood}{
\log \mathcal{L} \left( \mathcal{D} \mid \boldsymbol{\theta} \right) = \sum_{i=1}^N \log \ell_i
}\end{equation}

The methodology we use to infer the model's parameters is detailed in
section~\ref{sec-sampling}.

\section{Extension to a Bilateral Model}\label{sec-ext-to-contra}

A straightforward approach to modeling contralateral lymphatic spread
would be to use two independent unilateral models, as described in
Section section~\ref{sec-unilateral}, with shared parameters such as the
distribution of diagnosis times or spread between LNLs (\(t_{vr}\)).
However, this method would fail to capture the correlation between
ipsilateral and contralateral involvement discussed in
section~\ref{sec-data-strat}, particularly the observed increase in
contralateral involvement with greater severity of ipsilateral spread.

Thus, we extend the formalism in section~\ref{sec-unilateral} in such a
way that the model's ipsi- and contralateral side evolve synchronously
over time. To achieve that, we start by writing down the posterior
distribution of involvement in analogy to
equation~\ref{eq-uni-bayes-law}, which is now a joint probability of an
involvement \(\mathbf{X}^\text{i}\) ipsilaterally \emph{and} an
involvement \(\mathbf{X}^\text{c}\) contralaterally, given a diagnosis
of the ipsilateral LNLs \(\mathbf{Z}^\text{i}\) and of the contralateral
ones \(\mathbf{Z}^\text{c}\):

\begin{equation}\phantomsection\label{eq-bilateral-bayes}{
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right) = \frac{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right)}{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right)}
}\end{equation}

For the sake of brevity, we omit the dependency on the parameters and
the T-category here.

The probability of the diagnoses given a hidden state factorises:
\(P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) = P \left( \mathbf{Z}^\text{i} \mid \mathbf{X}^\text{i} \right) \cdot P \left( \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{c} \right)\),
and the two factors are described through observation matrices
\(\mathbf{B}^\text{i}\) and \(\mathbf{B}^\text{c}\).

The term representing the model's prior probability of hidden
involvement does not factorize. However, we assume no direct lymphatic
drainage from ipsilateral to contralateral LNLs, as major lymph vessels
do not cross the mid-sagittal plane. In the graphical model, this
translates to the absence of directed arcs between ipsilateral and
contralateral LNLs, implying that contralateral tumor spread occurs
solely via the primary tumor. We can thus write the joint probability
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right)\) as a
factorising sum:

\begin{equation}\phantomsection\label{eq-bilateral-marginal}{
\begin{aligned}
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \right) &= \sum_{t=0}^{t_\text{max}} P(t) \cdot P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid t \right) \\
&= \sum_{t=0}^{t_\text{max}} P(t) \cdot P \left( \mathbf{X}^\text{i} \mid t \right) \cdot P \left( \mathbf{X}^\text{c} \mid t \right)
\end{aligned}
}\end{equation}

This assumption is intuitive: since no major lymph vessels cross the
midline, the ipsilateral and contralateral sides of the lymphatic
network evolve independently over time. However, they are indirectly
coupled through time. For example, a joint state with severe
contralateral involvement and limited ipsilateral involvement is
improbable: Severe contralateral involvement typically occurs at later
time steps, when limited ipsilateral involvement is unlikely.

Using equation~\ref{eq-bilateral-marginal} along with
equation~\ref{eq-lambda-matrix}, we can write the above distribution
algebraically as a product:

\begin{equation}\phantomsection\label{eq-bilateral-marginal-algebra}{
P \left( \mathbf{X}^\text{i} = \boldsymbol{\xi}_n, \mathbf{X}^\text{c} = \boldsymbol{\xi}_m \right) = \left[ \boldsymbol{\Lambda}^\intercal_\text{i} \cdot \operatorname{diag} P(\mathbf{t}) \cdot \boldsymbol{\Lambda}_\text{c} \right]_{n,m}
}\end{equation}

\subsection{Parameter Symmetries}\label{sec-params-symmetry}

The matrices \(\boldsymbol{\Lambda}\text{i}\) and
\(\boldsymbol{\Lambda}\text{c}\) could, in principle, be parameterized
with entirely separate parameters, allowing ipsilateral and
contralateral spread rates to differ completely. However, we simplify
the parameter space by sharing parameters between the two sides, based
on the following three assumptions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Shared Graph Structure}: Both ipsilateral and contralateral
  spread are described by the same graph shown in
  figure~\ref{fig-full-graph}.
\item
  \textbf{Symmetric Spread Among LNLs}: The spread among LNLs is assumed
  to be the same on both sides, reflecting the symmetric structure of
  the lymphatic system. Consequently, the spread rates between nodes
  should also be symmetric. This is formalized as:\\
  \begin{equation}\phantomsection\label{eq-symmetries}{
    \begin{aligned}
    b_v^\text{c} &\neq b_v^\text{i} \\
    t_{rv}^\text{c} &= t_{rv}^\text{i}
    \end{aligned}
    }\end{equation} for all \(v \leq V\) and
  \(r \in \operatorname{pa}(v)\).
\item
  \textbf{Asymmetric Spread from Tumor}: Tumor spread to the
  contralateral side varies depending on whether the tumor crosses the
  mid-sagittal plane. When the tumor extends over the midline, its
  spread rate to the contralateral LNLs should be larger than if the
  tumor were lateralized, but still smaller than the tumor's spread to
  the ipsilateral side. To model this, we introduce a mixing parameter
  \(\alpha \in [0,1]\), representing the ``degree of asymmetry''
  effectively mixing these two extremes:\\
  \begin{equation}\phantomsection\label{eq-mixing}{
    b_v^{\text{c},\epsilon=\texttt{True}} = \alpha \cdot b_v^\text{i} + (1 - \alpha) \cdot b_v^{\text{c},\epsilon=\texttt{False}}
    }\end{equation} This parameterization results in three sets of rates
  for tumor spread to the LNLs: (1) the spread to ipsilateral LNLs
  \(b^\text{i}_v\), (2) the spread to contralateral LNLs as long as the
  tumor is lateralized \(b_v^{\text{c},\epsilon=\texttt{False}}\), (3)
  the spread to contralateral LNLs when the tumor crosses the midline
  \(b_v^{\text{c},\epsilon=\texttt{True}}\). Despite this additional
  complexity, these rates collectively account for only
  \(2 \cdot 2^V + 1\) parameters due to their coupling via \(\alpha\).
\end{enumerate}

The parameter space now becomes:

\begin{equation}\phantomsection\label{eq-bi-param-space}{
\boldsymbol{\theta} = \left( \left\{ b_v^\text{i} \right\}, \left\{ b_v^\text{c} \right\}, \alpha, \left\{ t_{vr} \right\}, p_\text{early}, p_\text{adv.} \right) \quad \text{with} \quad \genfrac{}{}{0pt}{2}{v\leq V}{r\in\operatorname{pa}(v)}
}\end{equation}

This results in less than a doubling of parameters compared to the
unilateral model. From these parameters, we construct three transition
matrices: the unchanged \(\mathbf{A}_\text{i}\) for the ipsilateral
side, \(\mathbf{A}_\text{c}^{\epsilon=\texttt{False}}\) for
contralateral progression while the tumor is lateralized, and
\(\mathbf{A}_\text{c}^{\epsilon=\texttt{True}}\) for cases where the
tumor crosses the mid-sagittal plane.

\subsection{Modelling Midline Extension}\label{sec-midline}

Most tumors crossing the midline at the time of diagnosis likely began
as lateralized tumors that grew over the midline at a later stage. As a
result, the transition matrix
\(\mathbf{A}_\text{c}^{\epsilon=\texttt{True}}\) applies only to a
subset of time-steps.

To account for this, we model the tumor's extension over the
mid-sagittal line as a binary random variable. A tumor starts as
lateralized, with a finite probability \(p_\epsilon\) at each time step
of crossing the midline. The overall probabilities of a patient having a
clearly lateralized tumor or one extending over the mid-sagittal line
after \(t\) time steps are then given by

\[
\begin{aligned}
P(\epsilon = \texttt{False} \mid t) &= (1 - p_\epsilon)^t \\
P(\epsilon = \texttt{True} \mid t) &= 1 - P(\epsilon = \texttt{False} \mid t)
\end{aligned}
\]

Using this, it is straightforward to write down the matrix of state
distributions for all time-steps, as in equation~\ref{eq-lambda-matrix}
covering the contralateral hidden state evolution:

\[
\boldsymbol{\Lambda}_\text{c}^{\epsilon=\texttt{False}} =
\left(
\begin{array}{r}
\boldsymbol{\pi}^\intercal \cdot \left( \mathbf{A}_\text{c}^{\epsilon=\texttt{False}} \right)^{0\phantom{t_\text{max}}} \\
(1-p_\epsilon) \cdot \boldsymbol{\pi}^\intercal \cdot \left( \mathbf{A}_\text{c}^{\epsilon=\texttt{False}} \right)^{1\phantom{t_\text{max}}} \\
\hfill \vdots \hfill \\
(1-p_\epsilon)^{t_\text{max}} \cdot \boldsymbol{\pi}^\intercal \cdot \left( \mathbf{A}_\text{c}^{\epsilon=\texttt{False}} \right)^{t_\text{max}\phantom{0}} \\
\end{array}
\right)
\]

where we used the transition matrix
\(\mathbf{A}_\text{c}^{\epsilon=\texttt{False}}\) that depends on the
base spread parameters \(b_v^{\text{c},\epsilon=\texttt{False}}\).

The case of midline extension is more complex: we already marginalize
over the exact time step when the tumor grows over the mid-sagittal
line. However, at the point of crossing, the contralateral transition
matrix must switch to the increased spread rates,
\(b_v^{\text{c}, \epsilon=\texttt{True}}\), as defined by the linear
mixing in equation~\ref{eq-mixing}. To correctly perform this
marginalization, we iteratively construct the joint distribution
\(P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid t \right)\).

We begin at \(t=0\), where all contralateral LNLs are healthy (i.e.,
\(\mathbf{X}_\text{c}=\boldsymbol{\xi}_0\)) and the tumor is lateralized
(\(\epsilon=\texttt{False}\)):

\[
P \left( \mathbf{X}^\text{c} = \boldsymbol{\xi}_0, \epsilon=\texttt{False} \mid t=0 \right) = 1
\]

while all other states have zero probability.

At some later time step \(t=\tau+1\), there are two scenarios to
marginalize over:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{The tumor was lateralized at \(t=\tau\) and grew over the
  midline at \(t=\tau+1\):}\\
  In this case, the probability of midline extension at \(t=\tau+1\) is
  \(p_\epsilon\). This probability weights the contralateral state
  distribution that had previously evolved without increased
  contralateral spread.
\item
  \textbf{The tumor had already crossed the midline before
  \(t=\tau\):}\\
  Here, the tumor remains in the midline-crossed state with probability
  1. To account for this scenario, we simply include the distribution
  \(P\left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid \tau \right)\)
  from the previous time step.
\end{enumerate}

Combining these scenarios leads to a recursive formulation:

\[
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid \tau + 1 \right) = \big[ p_\epsilon P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{False} \mid \tau \right) + P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid \tau \right) \big]^\top \cdot \mathbf{A}_\text{c}^{\epsilon=\texttt{True}}
\]

We can collect the iteratively computed distributions for the midline
extension case to define the matrix over the states given all
time-steps, in analogy to equation~\ref{eq-lambda-matrix}:

\[
\boldsymbol{\Lambda}_\text{c}^{\epsilon=\texttt{True}} = \begin{pmatrix}
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid 0 \right) \\
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid 1 \right) \\
\vdots \\
P \left( \mathbf{X}^\text{c}, \epsilon=\texttt{True} \mid t_\text{max} \right) \\
\end{pmatrix}
\]

Using this, we can again write the joint of ipsi- and contralateral
involvement - now also for the case of mid-sagittal extension -
algebraically as before in equation~\ref{eq-bilateral-marginal-algebra}:

\[
P \left( \mathbf{X}^\text{i} = \boldsymbol{\xi}_n, \mathbf{X}^\text{c} = \boldsymbol{\xi}_m, \epsilon \right) = \left[ \boldsymbol{\Lambda}^\intercal_\text{i} \cdot \operatorname{diag} P(\mathbf{t}) \cdot \boldsymbol{\Lambda}_\text{c}^\epsilon \right]_{n,m}
\]

With the above, we compute the likelihood of all patients with and
without midline extension separately. And if for some patients the
information of tumor lateralization is not available, we can simply
marginalize over the unknown variable
\(\epsilon \in \{ \texttt{False}, \texttt{True} \}\).

The final parameter space of our extended model has now reached this
size:

\begin{equation}\phantomsection\label{eq-ext-param-space}{
\boldsymbol{\theta} = \left( \left\{ b_v^\text{i} \right\}, \left\{ b_v^\text{c} \right\}, \alpha, \left\{ t_{vr} \right\}, p_\text{early}, p_\text{adv.}, p_\epsilon \right) \quad \text{with} \quad \genfrac{}{}{0pt}{2}{v\leq V}{r\in\operatorname{pa}(v)}
}\end{equation}

\subsection{Model Prediction in the Bayesian
Context}\label{model-prediction-in-the-bayesian-context}

Our stated goal is to compute the risk for a patient's true ipsi- and
contralateral nodal involvement states \(\mathbf{X}^\text{i}\) and
\(\mathbf{X}^\text{c}\), \emph{given} their individual diagnosis
\(d = \left( \boldsymbol{\zeta}^\text{i}_k, \boldsymbol{\zeta}^\text{c}_\ell, \epsilon, \text{T}x \right)\).
Here, this diagnosis consists of the observed ipsi- and contralateral
nodal involvements, the patient's midline extension \(\epsilon\), and
their tumor's T-category \(\text{T}x\). Using Bayes' law, we can write
this risk as:

\begin{equation}\phantomsection\label{eq-uni-bayes-law}{
P \big( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid d, \boldsymbol{\hat{\theta}} \big)
= \frac{P \left( \boldsymbol{\zeta}^\text{i}_k \mid \mathbf{X}^\text{i} \right) P \left( \boldsymbol{\zeta}^\text{c}_\ell \mid \mathbf{X}^\text{c} \right) P \big( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \mid \boldsymbol{\hat{\theta}}, \text{T}x \big)}
{\sum_{i=0}^{2^V} \sum_{j=0}^{2^V} \mathcal{C}_{ij}}
}\end{equation}

with the normalization constants

\[
\mathcal{C}_{ij} = P \left( \boldsymbol{\zeta}^\text{i}_k \mid \mathbf{X}^\text{i}=\boldsymbol{\xi}^\text{i}_i \right) P \big( \boldsymbol{\zeta}^\text{c}_\ell \mid \mathbf{X}^\text{c}=\boldsymbol{\xi}^\text{c}_j \big) P \big( \mathbf{X}^\text{i}=\boldsymbol{\xi}^\text{i}_i, \mathbf{X}^\text{c}=\boldsymbol{\xi}^\text{c}_j, \epsilon \mid \boldsymbol{\hat{\theta}}, \text{T}x \big)
\]

The terms
\(P \left( \boldsymbol{\zeta}^\text{i}_k \mid \mathbf{X}^\text{i} \right)\)
and
\(P \left( \boldsymbol{\zeta}^\text{c}_\ell \mid \mathbf{X}^\text{c} \right)\)
are defined solely by sensitivity and specificity of the diagnostic
modality. Terms like these already appeared in the definition of the
observation matrx in equation~\ref{eq-obs-matrix}. The \emph{prior}
\(P \big( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \mid \boldsymbol{\hat{\theta}}, \text{T}x \big)\)
in the above equation is the crucial term that is supplied by a trained
model and its parameters \(\boldsymbol{\hat{\theta}}\).

It is possible to compute this \emph{posterior} probability of true
involvement not only for one fully defined state
\((\mathbf{X}^\text{i}, \mathbf{X}^\text{c})\), but also for
e.g.~individual LNLs: For example, the risk for involvement in the
contralateral level IV would be a marginalization over all ipsilateral
states \(\boldsymbol{\xi}^\text{i}_i\) and all contralateral states
\(\boldsymbol{\xi}^\text{c}_j\) where \(\xi^\text{c}_{j4}=1\). Formally:

\begin{equation}\phantomsection\label{eq-marg-over-posterior}{
P \big( \text{cIV} \mid \mathbf{Z}^\text{i}=\boldsymbol{\zeta}^\text{i}_k, \mathbf{Z}^\text{c}=\boldsymbol{\zeta}^\text{c}_\ell, \boldsymbol{\hat{\theta}}, \text{T}x \big) = \sum_k \sum_{\ell \, : \, \xi_{\ell 4}=1} P \big( \mathbf{X}^\text{i} = \boldsymbol{\xi}^\text{i}_k, \mathbf{X}^\text{c} = \boldsymbol{\xi}^\text{c}_\ell \mid \boldsymbol{\zeta}^\text{i}_k, \boldsymbol{\zeta}^\text{c}_\ell, \epsilon, \boldsymbol{\hat{\theta}}, \text{T}x \big)
}\end{equation}

\section{Computational methods}\label{sec-methods}

This section details the experimental setup. All figures, tables, and
results are fully reproducible via the GitHub repository
\href{https://github.com/rmnldwg/bilateral-paper}{rmnldwg/bilateral-paper}.

\subsection{Involvement Data Consensus}\label{sec-methods-consensus}

Rather than providing the model with a patient's diagnoses from multiple
modalities with different sensitivities and specificities (e.g.~PET-CT
\emph{and} MRI), we combined them into a single ``consensus'' diagnosis
before inference. This approach was necessary because
literature-reported diagnostic imaging sensitivities and specificities
did not plausibly match our observations. For instance, in the USZ
cohort, 78\% of OPSCC patients were diagnosed with ipsilateral LNL II
involvement -\/--- a finding difficult to reconcile with typical imaging
sensitivities around 80\% and specificities below 100\%
\citep{debondt_detection_2007, kyzas_18ffluorodeoxyglucose_2008}.

\subsection{MCMC Sampling}\label{sec-sampling}

We used the Python package
\href{https://emcee.readthedocs.io/en/stable/}{\texttt{emcee}}
\citep{foreman-mackey_emcee_2013} for parameter inference, implementing
efficient MCMC sampling with parallel affine-invariant samplers. The
sampling algorithms employed differential evolution moves
\citep{terbraak_differential_2008, nelson_run_2013}, with the likelihood
implemented by our
\href{https://lymph-model.readthedocs.io/en/stable/}{\texttt{lymph-model}}
Python package.

We initialized 12 parallel samplers (``walkers'') with random values
from the unit cube, effectively representing a uniform prior
distribution over the model parameters. Convergence was determined by
two criteria:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The change in autocorrelation time was less than 5.0e-2.
\item
  The autocorrelation estimate dropped below \(n\) / 50, where \(n\) is
  the chain length. Earlier autocorrelation estimates might not be
  trustworthy.
\end{enumerate}

Samples from this \emph{burn-in phase} before convergence were
discarded. After that, we drew 10 additional samples, spaced 10 steps
apart.

We verified sampling convergence in
figure~\ref{fig-model-burnin-history} by examining the MCMC chain's
autocorrelation time and walker acceptance fractions.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_burnin_history.pdf}}

}

\caption{\label{fig-model-burnin-history}Burn-in phase monitoring of
MCMC sampling. Left: Estimated autocorrelation time, indicating
converging when stable and below the trust threshold. Right: Average
acceptance fraction of parallel walkers, with \textasciitilde30\%
indicating good mixing.}

\end{figure}%

\subsection{Computing the Observed and Predicted Prevalence of
Involvement Patterns}\label{sec-prevalence}

We evaluate the model's ability to approximate the observed distribution
of lymphatic involvement patterns. We compare the prevalence of selected
involvement patterns in the data to the model's predicted prevalence,
given patient scenarios.

A ``scenario'' includes the patient's T-category \(\text{T}x\) and
whether the tumor extended over the mid-sagittal line,
i.e.~\(\epsilon=\texttt{True}\) or \(\epsilon=\texttt{False}\). An
involvement pattern specifies all ipsi- and contralateral LNLs' status
as ``healthy'', ``involved'', or ``masked'' (ignored).

For example, we may assess contralateral LNL II involvement prevalence
under early T-category (T0-T2) and no midline extension
(\(\epsilon=\texttt{False}\)). In the data, 379 such patients were
observed, with 27 exhibiting contralateral LNL II involvement -- a
prevalence of 7.12\%.

To visualize the data prevalence, we plot a \emph{beta posterior} over
the true prevalence \(q\). It is the result of a uniform beta prior over
\(q\), multiplied with the binomial likelihood for \(k\) out of \(n\)
patients, given \(q\). The resulting distribution has its maximum at
\(q=k / n\) and nicely captures the uncertainty in the observed cohort:
if only 3 out of 10 patients are observed with a specific pattern, the
distribution is much wider than if we observe 300 out of 1000.

The model's predicted prevalence to compare it with is computed as:

\[
P \left( \text{II}^\text{c} \mid \epsilon=\texttt{False}, \text{T}x=\text{early} \right) = \frac{P \left( \text{II}^\text{c}, \epsilon=\texttt{False} \mid \text{T}x=\text{early} \right)}{P \left( \epsilon=\texttt{False} \mid \text{T}x=\text{early} \right)}
\]

In the enumator, we marginalize over all ipsi- and contralateral LNLs'
involvements, except for LNL II contralaterally. This is similar to the
marginalization in equation~\ref{eq-marg-over-posterior}, although we
are summing over different quantities. In the denominator, we can simply
insert the joint distribution over midline extension and diagnose time
\(P \left( \epsilon, t \right)\) marginalized over \(t\) using the early
T-category's time-prior.

We display model predictions as histograms, each value computed from one
of the MCMC samples. Ideally, these approximate the location and width
of the beta posteriors from the data showing an accurate and precise
fit.

Note that we omit the y-axis in these figures, as their numerical value
is not intuitively interpretable. We instead use the free space to label
e.g.~rows in an array of subplots.

\section{Results: Model evaluation}\label{sec-results}

In table~\ref{tbl-midline-params}, we tabulate the mean and standard
deviation of the sampled parameters. The bilateral model mostly
reproduces the ipsilateral spread parameter values reported in the
earlier publication on the unilateral model
\citep{ludwig_modelling_2023}. Any discrepancies may be due to
differences in the patient cohorts. Therefore, we omit the analysis of
the ipsilateral spread patterns and focus instead on the analysis of
contralateral involvement. Lastly, note the relatively small
contralateral spread parameters \(b^c_v\) compared to the ipsilateral
parameters \(b^i_v\). This accurately reflects the low prevalence of
contralateral lymph node involvement for lateralized tumors.

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\begin{longtable}[]{@{}lll@{}}

\caption{\label{tbl-midline-params}Mean sampled parameter estimates of
the midline model and the respective standard deviation. The parameters
set to fixed values are the maximum number of time steps \(t_{max}=10\)
and the time prior parameter for early T-category patients
\(p_{early}=0.3\).}

\tabularnewline

\caption{}\label{T_d0076}\tabularnewline
\toprule\noalign{}
Parameter & Mean & Std. Dev. \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Parameter & Mean & Std. Dev. \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Mid. ext. probability & 8.16\% & ± 0.48\% \\
ipsi: T ➜ I & 2.80\% & ± 0.26\% \\
ipsi: T ➜ II & 34.89\% & ± 1.40\% \\
ipsi: T ➜ III & 5.45\% & ± 0.66\% \\
ipsi: T ➜ IV & 0.94\% & ± 0.18\% \\
ipsi: T ➜ V & 1.83\% & ± 0.22\% \\
ipsi: T ➜ VII & 2.32\% & ± 0.26\% \\
contra: T ➜ I & 0.29\% & ± 0.09\% \\
contra: T ➜ II & 2.46\% & ± 0.29\% \\
contra: T ➜ III & 0.14\% & ± 0.07\% \\
contra: T ➜ IV & 0.19\% & ± 0.08\% \\
contra: T ➜ V & 0.05\% & ± 0.04\% \\
contra: T ➜ VII & 0.50\% & ± 0.17\% \\
Mixing ⍺ & 33.87\% & ± 4.32\% \\
I ➜ II & 62.50\% & ± 16.69\% \\
II ➜ III & 14.23\% & ± 1.64\% \\
III ➜ IV & 15.86\% & ± 1.93\% \\
IV ➜ V & 14.58\% & ± 3.76\% \\
late T-cat. binom. prob. & 44.98\% & ± 1.99\% \\

\end{longtable}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\subsection{Illustration of the model}\label{illustration-of-the-model}

In this subsection, we illustrate key aspects of the mathematical
framework introduced earlier. The top panel of
figure~\ref{fig-model-midext-evo} shows the prior distribution over
diagnosis times, \(P(t)\). Based on the parameterization, early
T-category tumors are on average diagnosed after 3 time steps, while
advanced T-category tumors are diagnosed later, averaging 4.5 time
steps. This is due to the learned value of \(p_\text{adv.}\), which is
44.98\% for advanced T-category tumors. The tumor's average probability
per time step of growing over the midline, \(p_\epsilon\), was found to
be 8.16\%. Using this value, the conditional probability of midline
extension, \(P(\epsilon \mid t)\), can be computed for a given time step
\(t\) (red line in the top panel of figure~\ref{fig-model-midext-evo}).
The bottom panel visualizes the joint probability \(P(\epsilon, t)\),
showing the likelihood of diagnosis at time \(t\) with specific states
of midline extension and T-category.

\begin{figure}

\centering{

\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/figures/fig_model_midext_evo.pdf}

}

\caption{\label{fig-model-midext-evo}The top panel shows the prior
probability to be diagnosed at time step \(t\) for early and late
T-category tumors as bars. The conditional probability of midline
extension (\(\epsilon=\texttt{True}\)) given time step \(t\) is shown as
a line plot. The bottom panel illustrates the joint probability of being
diagnosed at time \(t\) and having a tumor that crosses the midline.}

\end{figure}%

The framework models the joint probability distribution of midline
extension and ipsi- and contralateral lymph node involvement,
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right)\).
This is visualized in figure~\ref{fig-model-algebra-equation}, which
represents the calculation defined in
equation~\ref{eq-bilateral-marginal-algebra}. To simplify
interpretation, the example focuses only on LNLs II, III, and IV,
reducing the state space to \(2^3 = 16\) possible states per side, and
\(2 \times 8 \times 8 = 128\) total states. LNLs I, V, and VII are
excluded, along with their spread parameters, while remaining parameters
are set to their mean values from table~\ref{tbl-midline-params}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_algebra_equation.pdf}}

}

\caption{\label{fig-model-algebra-equation}Visual representation of
equation~\ref{eq-bilateral-marginal-algebra} for midline extension. The
left and right matrices represent the time evolution of hidden states
for the ipsi- and contralateral necks, respectively. The right matrices
distinguish between the cases of no midline extension (top) and midline
extension (bottom). The central diagonal matrix shows the time-prior for
late T-category tumors. This computation yields the joint distribution
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right)\),
visualized in figure~\ref{fig-model-state-dist}.}

\end{figure}%

The left matrix in figure~\ref{fig-model-algebra-equation} shows the
time evolution of the probability distribution over the ipsilateral
involvement states, starting from the healthy state \([0,0,0]\). The two
right matrices show the contralateral state evolution, distinguishing
between the cases of midline extension and no midline extension. At
\(t=0\), the contralateral neck begins in the healthy state \([0,0,0]\)
without midline extension. The central matrix shows the time prior for
late T-category tumors. The matrix multiplication results in the joint
distribution
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon=\texttt{True} \right)\),
visualized in figure~\ref{fig-model-state-dist}.

This joint distribution is presented as two heatmaps, corresponding to
the two states of midline extension. The most likely state involves a
lateralized tumor with ipsilateral level II involvement and no
contralateral involvement, having a probability of approximately 25\%.
The next most probable state involves ipsilateral levels II and III
without contralateral involvement. The most likely state with
contralateral involvement corresponds to tumors with midline extension,
showing involvement of contralateral level II and ipsilateral levels II
and III.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_state_dist.pdf}}

}

\caption{\label{fig-model-state-dist}The full distribution
\(P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon=\text{False} \right)\)
over ipsi- and contralateral states and midline extension for late
T-category tumors. This three-dimensional distribution is shown as two
separate heatmaps for the binary midline extension variable
\(\epsilon\). These matrices are the result of
equation~\ref{eq-bilateral-marginal-algebra}, visualized in
figure~\ref{fig-model-algebra-equation}.}

\end{figure}%

\subsection{Prevalence predictions for contralateral
involvement}\label{prevalence-predictions-for-contralateral-involvement}

The bilateral model was designed to meet the requirements outlined in
section~\ref{sec-requirements}. Here, we evaluate the model's ability to
quantitatively capture the observed patterns of lymph node involvement
in the dataset. Specifically, we compare the model's predictions for
contralateral involvement to the observed data across scenarios that
vary by T-category, midline extension, and ipsilateral involvement.

\subsubsection{Dependence of Contralateral Involvement on T-Category and
Midline
Extension}\label{dependence-of-contralateral-involvement-on-t-category-and-midline-extension}

In figure~\ref{fig-model-prevalences-overall}, we compare the prevalence
of contralateral involvement for LNLs II, III, and IV across four
scenarios: combinations of early and late T-category, and lateralized
versus midline-extending tumors.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_overall.pdf}}

}

\caption{\label{fig-model-prevalences-overall}Comparison of predicted
(histograms) vs observed (beta posteriors) prevalences, shown for the
contralateral LNLs II (blue), III (orange), and IV (green). The top row
shows scenarios with early T-category tumors, the bottom row for late
T-category ones. The left column depicts scenarios where the primary
tumor is clearly lateralized, the right column scenarios of tumors
extending over the mid-sagittal line. This figure illustrates the
model's ability to describe the prevalence of involvement for different
combinations of the risk factors T-category and midline extension.}

\end{figure}%

Figure~\ref{fig-model-prevalences-overall} demonstrates the model's
ability to account for key risk factors, such as T-category and midline
extension. Consistent with the data, the model predicts that the
prevalence of contralateral LNL II involvement increases from 7.1\% for
early T-category lateralized tumors to 39.17\% for advanced T-category
tumors that cross the midline. Similarly, contralateral LNL III
involvement rises from around 1.55\% for early T-category lateralized
tumors to nearly 14.25\% for advanced T-category midline-extending
tumors.

\subsubsection{Influence of Upstream Involvement on Contralateral
Metastasis}\label{influence-of-upstream-involvement-on-contralateral-metastasis}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_upstream.pdf}}

}

\caption{\label{fig-model-prevalences-upstream}The influence of the
upstream LNL II's involvement on the prevalence of contralateral level
III for the four combinations of tumor lateralization (lateralized or
extending over midline) and T-category (early or advanced). Our model
predictions (histograms) are plotted against the observations in the
data (beta posteriors).}

\end{figure}%

Figure~\ref{fig-model-prevalences-upstream} highlights the influence of
upstream LNL II involvement on contralateral LNL III metastasis. The
contralateral LNL III rarely harbors metastases when its upstream LNL II
is healthy, a correlation well-captured by the model. However, for early
T-category tumors extending over the midline, the model slightly
underestimates the observed prevalence, likely due to the relative
rarity of such cases, resulting in broader uncertainty in the observed
prevalence estimates.

\subsubsection{Correlation between Ipsi- and Contralateral
Involvement}\label{correlation-between-ipsi--and-contralateral-involvement}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_with_ipsi.pdf}}

}

\caption{\label{fig-model-prevalences-with-ipsi}Comparison of the
computed and observed prevalences for scenarios that illustrate the
model's capability of accounting for the correlation between ipsi- and
contralateral involvement. We show three scenarios where we consider the
joint involvement of contralateral LNL II together with different
ipsilateral involvements: 1) the ipsilateral neck shows no involvement
in blue (LNLs I to V are healthy, LNL VII is unspecified because data on
it is missing for some patients), 2) where ipsilateral LNL II is
involved in orange (LNLs I, III, IV, and V are healthy), and 3) where
ipsilateral LNLs II and III are involved in orange (LNLs I, IV, and V
are healthy). These two scenarios are plotted for all combinations of
T-category (early in top row, advanced in bottom row) and tumor
lateralization (lateralized in left column, extending over mid-sagittal
line in the right column).}

\end{figure}%

In figure~\ref{fig-model-prevalences-with-ipsi}, the model's ability to
capture the correlation between ipsi- and contralateral involvement is
demonstrated. The marginals of the joint distribution highlight
contralateral LNL II involvement alongside varying ipsilateral LNL
involvement states. Despite having no direct connections between the two
sides, the model successfully predicts these correlations, which arise
purely through the shared diagnosis time.

For example, the model accurately predicts that contralateral LNL II
involvement is rare when the ipsilateral neck is completely healthy
(green histograms). However, if ipsilateral LNL II is involved,
contralateral involvement becomes more likely. Notably, the model
achieves this without specific parameters to quantify ipsi-contralateral
correlations, relying instead on the inherent structure of the
time-dependent dynamics.

\section{Results: Prediction of Risk for Occult
Disease}\label{sec-results-risk}

For clinical applications, the model needs to estimate the risk of
occult metastases in clinically negative LNLs based on a patient's
diagnosis. The diagnosis includes the T-category, tumor lateralization
(whether it extends over the mid-sagittal plane), and the clinically
detected involvement of LNLs based on imaging and possibly fine needle
aspiration (FNA).

We assume imaging detects lymph node involvement with a sensitivity of
81\% and a specificity of 76\% \citep{debondt_detection_2007}, while FNA
has a sensitivity of 80\% and a specificity of 98\%. This implies that
FNA-confirmed involvement is highly reliable, with almost no false
positives.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/mediabag/figures/fig_model_risks.pdf}}

}

\caption{\label{fig-model-risks}Histograms over the predicted risk of
occult involvement in contralateral LNL II (left), III (middle), and IV
(right), shown for various combinations of T-category, tumor
lateralization, and clinical LNL diagnoses. All LNLs not explicitly
mentioned in the legend, including the LNL for which the risk of occult
disease was computed, were assumed to be clinically negative
(specificity 76\%, sensitivity 81\%).}

\end{figure}%

\subsection{Contralateral LNL II}\label{contralateral-lnl-ii}

The predicted risk of occult disease in contralateral LNL II is shown in
the left panel of figure~\ref{fig-model-risks}. Tumor lateralization is
the strongest determinant of risk. For example:

\begin{itemize}
\tightlist
\item
  A patient with a lateralized early T-category tumor and ipsilateral
  LNL II involvement has a predicted risk of 1.55\% for occult
  contralateral LNL II disease (green histogram).\\
\item
  If the same tumor extends over the midline, the risk increases to
  7.58\% (orange histogram).
\end{itemize}

Advanced T-category further increases risk but has less impact than
midline extension. For instance:

\begin{itemize}
\tightlist
\item
  An early T-category tumor crossing the midline with ipsilateral
  involvement of LNLs II and III has a 9.06\% risk of contralateral LNL
  II disease (red histogram).\\
\item
  For the same scenario with an advanced T-category tumor, the risk
  rises to 11.25\% (purple histogram).
\end{itemize}

The degree of ipsilateral involvement also influences risk. For a
midline-extending early T-category tumor, a clinically N0 ipsilateral
neck predicts a risk of 7.58\%, compared to 9.06\% when LNLs II and III
are involved.

In summary, midline extension is the primary risk factor for
contralateral LNL II involvement, but advanced T-category and extensive
ipsilateral involvement also contribute significantly.

\subsection{Contralateral LNL III}\label{contralateral-lnl-iii}

As shown in the center panel of figure~\ref{fig-model-risks}, the risk
of occult contralateral LNL III involvement rarely exceeds 5\% and
depends strongly on upstream LNL II involvement:

\begin{itemize}
\tightlist
\item
  For an advanced T-category tumor extending over the midline and with
  extensive clinical involvement in the ipsilateral LNLs II, III, and
  IV, but a clinically negative contralateral neck, the risk is only
  2.07\% (green histogram).\\
\item
  If contralateral LNL II is clinically involved, the risk for
  contralateral LNL III rises to 4.77\% (blue histogram).\\
\item
  When contralateral LNL II involvement is confirmed by FNA, the risk
  further increases to 6.94\% (red histogram).
\end{itemize}

Even for lateralized tumors, FNA-confirmed involvement in contralateral
LNL II predicts a 5.58\% risk for LNL III involvement (orange
histogram). This highlights the importance of upstream LNL II
involvement in determining the risk for LNL III.

\subsection{Contralateral LNL IV}\label{contralateral-lnl-iv}

For contralateral LNL IV, the predicted risk is below 1\% in most
scenarios, even for advanced T-category tumors extending over the
midline with extensive ipsilateral and contralateral involvement (green
histogram).

\begin{itemize}
\tightlist
\item
  If contralateral LNL III is also clinically involved, the risk
  increases to 2.81\% (blue histogram).\\
\item
  When contralateral LNL III involvement is confirmed by FNA, the risk
  rises significantly to 5.68\% (orange histogram).
\end{itemize}

This higher risk occurs because FNA confirmation eliminates the
possibility of false-positive diagnoses for contralateral LNL III,
strongly increasing the likelihood of downstream LNL IV involvement.

\subsection{Contralateral LNLs I, V, and
VII}\label{contralateral-lnls-i-v-and-vii}

Contralateral LNLs I, V, and VII show very low predicted risks for
occult disease. Even in advanced T-category tumors extending over the
midline with extensive ipsi- and contralateral involvement, the risk for
LNLs I and VII remains below 3\%.

For contralateral LNL V, the risk is negligible unless there is severe
contralateral involvement, including LNL IV, confirmed by FNA. In such
extreme cases, the risk increases to 5.49\%.

\section{Discussion}\label{sec-discussion}

\subsection{Summary}\label{summary}

This work introduces a formalism to model ipsi- and contralateral lymph
node involvement in oropharyngeal SCC patients. Building on a previously
developed ipsilateral model
\citep{ludwig_hidden_2021, ludwig_modelling_2023}, we extend it to the
contralateral side while preserving the original model's structure. Our
extension is both intuitive and interpretable, with parameters learned
from a dataset of 833 patients across four institutions. The model uses
clinically diagnosed LNL involvement, tumor T-category, and
lateralization to provide personalized risk predictions for occult
disease in any LNL of interest.

Despite its relatively few parameters, the model is highly
interpretable, with each parameter having a clear, intuitive
explanation. At the same time, it effectively describes the observed
data on ipsilateral and contralateral nodal involvement. To our
knowledge, this represents the most comprehensive model of lymphatic
tumor progression in oropharyngeal SCC, surpassing prior efforts, which
were conceptually different, limited in scope, or not trained on real
patient data \citep{benson_markov_2006, jung_development_2017}. The
underlying dataset and code are publicly available, supporting
reproducibility and further development.

\subsection{Implications for Contralateral Elective Nodal
Treatment}\label{implications-for-contralateral-elective-nodal-treatment}

The model's predictions are already guiding a clinical trial on volume
de-escalation at the University Hospital Zurich
\citep{universityofzurich_personalized_2024}. Based on a 5\% acceptable
risk threshold for occult disease in any LNL, the model recommends the
following approach for contralateral elective irradiation, assuming the
respective LNL is clinically healthy:

\begin{itemize}
\tightlist
\item
  \textbf{Lateralized tumors with no contralateral clinical
  involvement:}\\
  Unilateral radiotherapy is sufficient, regardless of T-category or
  ipsilateral involvement.\\
\item
  \textbf{Tumors extending over the midline with no contralateral
  clinical involvement:}\\
  Elective irradiation is limited to LNL II.\\
\item
  \textbf{Contralateral LNL III:}\\
  Irradiate only if LNL II is involved, regardless of tumor
  lateralization, T-category, or ipsilateral involvement. If
  contralateral LNL II is clinically negative, LNL III is not irradiated
  unless LNL IV is involved.\\
\item
  \textbf{Contralateral LNL IV:}\\
  Irradiate only when LNL III involvement is confirmed.\\
\item
  \textbf{Contralateral LNL V:}\\
  Elective irradiation is generally unnecessary. Only in extreme cases,
  such as advanced T-category tumors with midline extension and
  confirmed contralateral involvement in LNLs II to IV, is irradiation
  of LNL V considered.\\
\item
  \textbf{Contralateral LNLs I and VII:}\\
  Elective irradiation is not recommended unless these nodes are
  clinically involved.
\end{itemize}

These recommendations align with the model results discussed in
section~\ref{sec-results-risk}, providing a basis for refined treatment
guidelines. However, they should be interpreted in light of the
limitations discussed in section~\ref{sec-discussion-limitations} below.

\subsection{Limitations and Future
Work}\label{sec-discussion-limitations}

\subsubsection{T-Category Dependence}\label{t-category-dependence}

The model uses a single parameter, \(p_\text{adv.}\), to account for
differences in lymph node involvement patterns between early and
advanced T-category tumors. Advanced T-category tumors are modeled as
evolving over more time steps, while the probability of spread per time
step, governed by the \(b_v\) parameters, remains constant. While this
approach captures the overall differences between early and advanced
T-category tumors well (as seen in
figure~\ref{fig-model-prevalences-overall}), it is not perfect.

For example: - The observed differences between early and advanced
T-category tumors may sometimes be greater or smaller than the model's
predictions. This has been previously noted for ipsilateral LNL I
involvement \citep{ludwig_modelling_2023}. - For contralateral spread,
the prevalence of midline extension is slightly overestimated for early
T-category tumors and underestimated for advanced T-category tumors, as
discussed in section~\ref{sec-prevalence-midext}.

\subsubsection{Sensitivity and
Specificity}\label{sensitivity-and-specificity}

As noted in section~\ref{sec-methods-consensus} and further detailed in
section~\ref{sec-consensus}, we assumed that the consensus across
diagnostic modalities reflects the true state \(X_v\) of lymph node
involvement. While this assumption is reasonable for pathologically
confirmed diagnoses, it is an approximation for clinically diagnosed
involvement, which cannot detect occult disease by definition.

The model could, in principle, distinguish between pathologically
confirmed and clinically diagnosed involvement by incorporating
different sensitivity and specificity values for each diagnostic
modality. However, this was not implemented in the current work for two
reasons: 1. \textbf{Simplified evaluation:} This approximation allowed
for direct comparison between the observed prevalence of involvement and
the model's predictions, enabling an evaluation of the model's ability
to describe the data with few interpretable parameters. 2.
\textbf{Inconsistent literature values:} Reported sensitivity and
specificity values for diagnostic modalities show inconsistencies with
some observed data patterns, complicating their integration into the
model.

Future efforts could focus on developing methods to rigorously
differentiate between pathologically confirmed and clinically diagnosed
involvement, potentially improving model accuracy for clinical settings.

\subsubsection{Tumor Subsites}\label{tumor-subsites}

Oropharyngeal tumors occur in distinct subsites, such as the base of the
tongue or the tonsils, which may exhibit slightly different lymphatic
metastatic patterns. Incorporating subsite-specific information into the
model could enhance its predictive accuracy. Preliminary investigations
suggest that a mixture model may effectively capture these
subsite-specific spread patterns
\citep{sarrut_proceedings_2024, ludwig_modelling_2023a}.

This approach could also facilitate extending the model to other tumor
locations, such as the oral cavity, hypopharynx, and larynx. Including
these additional tumor sites would broaden the model's applicability and
improve its clinical utility for head and neck cancer patients.

\section{Acknowledgement}\label{acknowledgement}

This work was supported by:

\begin{itemize}
\tightlist
\item
  the Clinical Research Priority Program ``Artificial Intelligence in
  Oncological Imaging'' of the University of Zurich
\item
  the Swiss Cancer Research Foundation under grant number KFS
  5645-08-2022
\end{itemize}

\section{Consensus on most likely involvement}\label{sec-consensus}

The consensus on the most likely involvement state of a LNL was formed
as follows: Suppose the involvement status \(X_v\) of LNL \(v\) was
assessed using different diagnostic modalities
\(\mathcal{O} = \{ \text{MRI}, \text{CT}, \text{pathology}, \ldots \}\),
each characterized by their own pair of sensitivity and specificity
values \(s_N^{\mathcal{o}}\) and \(s_P^{\mathcal{o}}\), with
\(\mathcal{o} \in \mathcal{O}\). These values are tabulated in
table~\ref{tbl-spec-sens}. Then we have \(|\mathcal{O}|\) observations
\(z_v^{\mathcal{o}} \in \left[ 0, 1 \right]\), where 0 stands for
``healthy'' and 1 for ``involved''. We can then compute the most likely
true involvement \(X_v\) using the likelihood function

\[
\begin{aligned}
\ell \left( X_v \mid \{ z_v^{\mathcal{o}} \}_{\mathcal{o} \in \mathcal{O}} \right) = \prod_{\mathcal{o} \in \mathcal{O}}
\left( 1 - X_v \right) \cdot &\left[ z_v^{\mathcal{o}} \cdot \left( 1 - s_P^{\mathcal{o}} \right) + \left( 1 - z_v^{\mathcal{o}} \right) \cdot s_P^{\mathcal{o}} \right] \\
+ X_v \cdot &\left[ z_v^{\mathcal{o}} \cdot s_N^{\mathcal{o}} + \left( 1 - z_v^{\mathcal{o}} \right) \cdot (1 - s_N^{\mathcal{o}}) \right]
\end{aligned}
\]

We now assume the true state \(X_v\) to take on the value 1 if
\(\ell \left( X_v = 1 \mid \ldots \right) > \ell \left( X_v = 0 \mid \ldots \right)\)
and 0 otherwise. For example, if we have \(z_\text{II}^\text{CT} = 0\)
and \(z_\text{II}^\text{MRI} = 1\) we would compute the following
likelihoods:

\[
\begin{aligned}
\ell \left( X_\text{II} = 1 \mid z_\text{II}^\text{CT} = 0, z_\text{II}^\text{MRI} = 1 \right) &= \left( 1 - s_N^\text{CT} \right) \cdot s_N^\text{MRI} = 15.39\% \\
\ell \left( X_\text{II} = 0 \mid z_\text{II}^\text{CT} = 0, z_\text{II}^\text{MRI} = 1 \right) &= s_P^\text{CT} \cdot \left(1 - s_N^\text{MRI}\right) = 14.44\%
\end{aligned}
\]

In this example, we would thus assume the true state to be involved
(\(X_\text{II} = 1\)).

This method of computing a consensus also ensures that the pathology
reports always override any conflicting clinical diagnosis, due to
pathology's high sensitivity and specificity.

\begin{longtable}[]{@{}lrr@{}}
\caption{Specificity and sensitivity values from the literature
\citep{debondt_detection_2007, kyzas_18ffluorodeoxyglucose_2008}.}\label{tbl-spec-sens}\tabularnewline
\toprule\noalign{}
Modality & Specificity & Sensitivity \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Modality & Specificity & Sensitivity \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
CT & 76\% & 81\% \\
PET & 86\% & 79\% \\
MRI & 63\% & 81\% \\
FNA & 98\% & 80\% \\
pathology & 100\% & 100\% \\
\end{longtable}

\section{Contralateral Prevalence of
Involvement}\label{contralateral-prevalence-of-involvement}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\begin{longtable}[]{@{}llllllllllll@{}}

\caption{\label{tbl-data-strat}Contralateral involvement depending on
whether the primary tumor extends over the mid-sagittal line, the
T-category, and whether the ipsilateral LNL III was involved or
healthy.}

\tabularnewline

\caption{}\label{T_7a9be}\tabularnewline
\toprule\noalign{}
T-cat. & ipsi & Mid. ext. & \multicolumn{2}{l}{%
I} & \multicolumn{2}{l}{%
II} & \multicolumn{2}{l}{%
III} & \multicolumn{2}{l}{%
IV} & total \\
& & & n & \% & n & \% & n & \% & n & \% & n \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
T-cat. & ipsi & Mid. ext. & \multicolumn{2}{l}{%
I} & \multicolumn{2}{l}{%
II} & \multicolumn{2}{l}{%
III} & \multicolumn{2}{l}{%
IV} & total \\
& & & n & \% & n & \% & n & \% & n & \% & n \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
early & 0 & False & 0 & 1 & 0 & 0 & 0.00 & 1.16 & 0.00 & 0.00 & 86 \\
early & 0 & True & 0 & 1 & 1 & 0 & 0.00 & 10.00 & 10.00 & 0.00 & 10 \\
early & 0 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 12 \\
early & 1 & False & 1 & 11 & 2 & 1 & 0.53 & 5.82 & 1.06 & 0.53 & 189 \\
early & 1 & True & 1 & 2 & 0 & 0 & 11.11 & 22.22 & 0.00 & 0.00 & 9 \\
early & 1 & nan & 0 & 3 & 0 & 1 & 0.00 & 9.68 & 0.00 & 3.23 & 31 \\
early & ≥ 2 & False & 1 & 15 & 3 & 3 & 0.96 & 14.42 & 2.88 & 2.88 &
104 \\
early & ≥ 2 & True & 0 & 3 & 4 & 1 & 0.00 & 30.00 & 40.00 & 10.00 &
10 \\
early & ≥ 2 & nan & 0 & 3 & 0 & 0 & 0.00 & 13.64 & 0.00 & 0.00 & 22 \\
advanced & 0 & False & 0 & 2 & 0 & 0 & 0.00 & 5.88 & 0.00 & 0.00 & 34 \\
advanced & 0 & True & 0 & 3 & 0 & 0 & 0.00 & 12.50 & 0.00 & 0.00 & 24 \\
advanced & 0 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 3 \\
advanced & 1 & False & 0 & 3 & 0 & 0 & 0.00 & 4.55 & 0.00 & 0.00 & 66 \\
advanced & 1 & True & 1 & 18 & 5 & 1 & 1.64 & 29.51 & 8.20 & 1.64 &
61 \\
advanced & 1 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 9 \\
advanced & ≥ 2 & False & 4 & 16 & 7 & 4 & 5.26 & 21.05 & 9.21 & 5.26 &
76 \\
advanced & ≥ 2 & True & 3 & 46 & 18 & 8 & 3.80 & 58.23 & 22.78 & 10.13 &
79 \\
advanced & ≥ 2 & nan & 0 & 0 & 0 & 0 & 0.00 & 0.00 & 0.00 & 0.00 & 8 \\

\end{longtable}

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

\section{Prevalence of Midline Extension}\label{sec-prevalence-midext}

\begin{figure}

\centering{

\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/figures/fig_model_prevalences_midext.pdf}

}

\caption{\label{fig-model-prevalences-midext}Comparing the predicted
(histograms) and observed (lines depicting beta posteriors) prevalence
of midline extension for early (blue) and late (orange) T-category.
While the prevalence is predicted correctly when marginalizing over
T-category, the model cannot capture the degree of separation observed
in the data. Since the tumor's midline extension is virtually always
part of the diagnosis and hence \emph{given} when predicting a patient's
risk, we do not consider this discrepancy a major issue.}

\end{figure}%

In figure~\ref{fig-model-prevalences-midext}, we plot the prevalence of
midline extension in the data versus our model's prediction. It is
obvious the model cannot match the large spread between early and
advanced T-category seen in the data. This is because to achieve that,
it would need to increase the advanced T-category patient's prior
distribution over diagnosis times and at the same time reduce the
probability of the tumor to cross the midline during a time-step. But
since the time-priors parameter is also coupled with the spread
probabilities among the LNLs, the model does not have that freedom.

However, we do not consider this discrepancy a major limitation of the
model: We will not realistically be interested in the probability of
midline extension, as it is always possible to assess it with high
certainty. That is also the reason why we initially modelled the midline
extension \emph{not} as a random variable, but as a global risk factor
that would have been turned on or off from the onset of a patient's
disease evolution. This, however, lead to overly high risks for
contralateral involvement in advanced T-category patients with midline
extension, because then the model assumes an increased spread to the
contralateral side from the onset of the disease. Which is probably not
true in a majority of those cases. Thus, treating it as a random
variable that only becomes true during a patient's disease evolution
resulted in a better description of the data.

Formally, the wrong prediction of midline extension prevalence makes
little difference, since it is always given: Instead of
\(P\left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \right)\),
we typically compute
\(P\left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c}, \epsilon \right)\),
which does not suffer from the wrong probability of midline extension,
as the distribution over hidden states is renormalized:

\[
P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c} \mid \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c}, \epsilon \right) = \frac{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c} \mid \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right) P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon \right)}{P \left( \mathbf{Z}^\text{i}, \mathbf{Z}^\text{c}, \epsilon \right)}
\]

Note that a distribution over \(\epsilon\) appears both in the
enumerator and the denominator, which largely cancel each other, leaving
only the midline extension's effect on the distribution over hidden
states in the prediction.

Also, the discrepancy in midline extension prevalence between early and
advanced T-category is particularly pronounced in oropharyngeal SCC
patients. For example, in oral cavity SCC, the midline extension only
increases from 15.4\% (20 out of 130) to 33.3\% (13 out of 39).

\textsubscript{Source:
\href{https://rmnldwg.github.io/bilateral-paper/manuscript.qmd.html}{Article
Notebook}}

  \bibliography{references.bib}

\end{document}